---
title: "BioClip_analysis"
author: "Alma Steireif"
date: "`r Sys.Date()`"
output: html_document
---

Author: Alma Steireif
Date: `r Sys.Date()`

### Description

The goal of this script is to reformat the data obtained from the BioClip classification model. 

**flower_cam**: images that were taken above flowers
  * these are full frame pictures taken above a flower every 5 second   
  * they have gone through an object detection model to locate insects and crop a new image around them    
      + yolo11s model was used to detect insects    
  * the cropped images of the insects that have gone through a classification model to determine to the *family* level    
      + the classification tool used was pyBioClip (version: )
      + https://imageomics.github.io/bioclip/ , https://imageomics.github.io/pybioclip/    
      + Here's the output format of the classification model:    
      
| Image_Path          | Family                      | Family_Confidence                         | Classification_Category                                          |
|:--------------------|:----------------------------|:------------------------------------------|:-----------------------------------------------------------------|
| /path/to/image.jpg  | family predicted by bioclip | bioclip's confidence in the output class  | Category: stored only if ID belongs to one of 384 valid families |
| /path/to/image1.jpg | Coccinellidae               | 0.9778                                    | Coccinellidae                                                    |
| /path/to/image2.jpg | Bromeliaceae                | 0.1797                                    | other_families                                                   |
 

**platform_cam**: images taken above platforms   
  * these images were cropped on site by an object detector contained in the mini-computer accompanying the camera
  * these crops have gone through two classifiers: 
  * the first classifier is called InsectDetect, which used a YOLOv5 model to detect insects 
      + https://maxsitt.github.io/insect-detect-docs/, https://github.com/maxsitt/insect-detect
      + this model gives the bounding box of the insect and the confidence of the top1 class
  * the second classifier was pyBioClip again, with the same version as above. 
  * overall the results were combined to get this type of output: 
  
  | img_name | top1 | top1_prob | Family_BioClip | Family_Confidence_BioClip | Classification_Category_BioClip |
  |:---------|:-----|:----------|:---------------|:--------------------------|:--------------------------------|
  | img1.jpg | fly  | 0.28      | Muscidae       | 0.5115520358085632        | Muscidae                        |
  | img2.jpg | fly_small| 0.89  | Asilidae       | 0.7667312026023865        | Asilidae                        |
  
  For now I will put the platform_cam data aside and focus on the flower_cam data.

### Load pacakges and data

```{r setup,include=FALSE}
library(tidyverse)

sessionInfo()
```

### Flower data

```{r load data, message=FALSE, warning=FALSE}
# load the flower data for each camera
flower_cam31 <- read_csv("C:/Users/Almas/bioclip_test/BioClip_testrun/classifications_cam31.csv")
flower_cam32 <- read_csv("C:/Users/Almas/bioclip_test/BioClip_testrun/classifications_cam32.csv")
flower_cam33 <- read_csv("C:/Users/Almas/bioclip_test/BioClip_testrun/classifications_cam33.csv")
flower_cam34 <- read_csv("C:/Users/Almas/bioclip_test/BioClip_testrun/classifications_cam34.csv")
flower_cam35 <- read_csv("C:/Users/Almas/bioclip_test/BioClip_testrun/classifications_cam35.csv")
```

```{r filename information}
#create a function that retrieves data from the filename 
# this function will be used to create a new column with the camera number, one with the date and one with the time
get_info <- function(data){
  data %>%
    mutate(cam = str_extract(Image_Path, "cam\\d{2}"),
           date = str_extract(Image_Path, "\\d{4}-\\d{2}-\\d{2}"),
           #the time is found after the date and an undescore: 2021-06-01_12-00-00
           time = str_extract(Image_Path, "_(\\d{2}-\\d{2}-\\d{2})")) %>%
    #remove underscores from the time
    mutate(time = str_replace_all(time, "_", ""))
}

flower_cam31 <- get_info(flower_cam31)
flower_cam32 <- get_info(flower_cam32)
flower_cam33 <- get_info(flower_cam33)
flower_cam34 <- get_info(flower_cam34)
flower_cam35 <- get_info(flower_cam35)
```


```{r combine data}
#combine the data from all cameras
flower_cam <- bind_rows(flower_cam31, flower_cam32, flower_cam33, flower_cam34, flower_cam35)
```

```{r additional field data - site assignment}
flower_cam <- flower_cam %>%
mutate(site = case_when(
    grepl("2024-07-11", date) ~ "WED",  # Match the 11th July to Wedringen
    grepl("2024-07-14", date) ~ "STP",  # Match the 14th July to Steutz
    grepl("2024-07-18", date) ~ "JEP",  # Match the 18th July to Jerichow
    grepl("2024-07-20", date) ~ "KOT",  # Match the 20th July to Köthen
    grepl("2024-07-21", date) ~ "DES",  # Match the 21nd July to Dessau
    grepl("2024-07-24", date) ~ "HLI",  # Match the 24th July to Pölsfeld
    grepl("2024-07-29", date) ~ "WUP",  # Match the 29th July to Wulfen
    grepl("2024-07-30", date) ~ "WDG",  # Match the 30th July to Wedringen
    grepl("2024-08-13", date) ~ "BUH",  # Match the 13th August to Bühne
    TRUE ~ "UNKNOWN"  # Default value
  ))
```


```{r additional field data - flower assignment}
# Assign flower species to each camera trap
flower_cam <- flower_cam |>
  mutate(flower_sp = case_when(
    grepl("cam31", cam) & grepl("2024-07-11", date) ~ "Knautia arvensis", #cam31, 11th July, Knautia arvensis
    grepl("cam31", cam) & grepl("2024-07-14", date) ~ "Daucus carota", #cam31, 14th July, Daucus carota
    grepl("cam31", cam) & grepl("2024-07-18", date) ~ "Hypericum perforatum", #cam31, 18th July, Hypericum perforatum
    grepl("cam31", cam) & grepl("2024-07-20", date) ~ "Pastinaca sativa", #cam31, 20th July, Pastinaca sativa
    grepl("cam31", cam) & grepl("2024-07-21", date) ~ "Daucus carota", #cam31, 21st July, Daucus carota
    grepl("cam31", cam) & grepl("2024-07-24", date) ~ "Pastinaca sativa", #cam31, 24th July, Pastinaca sativa  
    grepl("cam31", cam) & grepl("2024-07-29", date) ~ "Pastinaca sativa", #cam31, 29th July, Pastinaca sativa
    grepl("cam31", cam) & grepl("2024-07-30", date) ~ "Galium album", #cam31, 30th July, Galium album
    grepl("cam31", cam) & grepl("2024-08-13", date) ~ "Lotus corniculatus", #cam31, 13th August, Lotus corniculatus
    
    grepl("cam32", cam) & grepl("2024-07-11", date) ~ "Knautia arvensis", #cam32, 11th July, Knautia arvensis
    grepl("cam32", cam) & grepl("2024-07-14", date) ~ "Crepis biennis", #cam32, 14th July, Crepis biennis
    grepl("cam32", cam) & grepl("2024-07-18", date) ~ "Peucedanum officinale", #cam32, 18th July, Peucedanum officinale
    grepl("cam32", cam) & grepl("2024-07-20", date) ~ "Knautia arvensis", #cam32, 20th July, Knautia arvensis
    grepl("cam32", cam) & grepl("2024-07-21", date) ~ "Crepis biennis", #cam32, 21st July, Crepis biennis
    grepl("cam32", cam) & grepl("2024-07-24", date) ~ "Knautia arvensis", #cam32, 24th July, Knautia arvensis
    grepl("cam32", cam) & grepl("2024-07-29", date) ~ "Crepis biennis", #cam32, 29th July, Crepis biennis
    grepl("cam32", cam) & grepl("2024-07-30", date) ~ "Pastinaca sativa", #cam32, 30th July, Pastinaca sativa
    grepl("cam32", cam) & grepl("2024-08-13", date) ~ "Galium album", #cam32, 13th August, Galium album
    
    grepl("cam33", cam) & grepl("2024-07-11", date) ~ "Daucus carota", #cam33, 11th July, Daucus carota
    grepl("cam33", cam) & grepl("2024-07-14", date) ~ "Trifolium repens", #cam33, 14th July, Trifolium repens
    grepl("cam33", cam) & grepl("2024-07-18", date) ~ "Galium boreale", #cam33, 18th July, Galium boreale
    grepl("cam33", cam) & grepl("2024-07-20", date) ~ "Crepis biennis", #cam33, 20th July, Crepis biennis
    grepl("cam33", cam) & grepl("2024-07-21", date) ~ "Crepis biennis", #cam33, 21st July, Crepis biennis
    grepl("cam33", cam) & grepl("2024-07-24", date) ~ "Hypericum perforatum", #cam33, 24th July, Hypericum perforatum
    grepl("cam33", cam) & grepl("2024-07-29", date) ~ "Pastinaca sativa", #cam33, 29th July, Pastinaca sativa
    grepl("cam33", cam) & grepl("2024-07-30", date) ~ "Centaurea jacea", #cam33, 30th July, Centaurea jacea
    grepl("cam33", cam) & grepl("2024-08-13", date) ~ "Trifolium pratense", #cam33, 13th August, Trifolium pratense
    
    grepl("cam34", cam) & grepl("2024-07-11", date) ~ "Pastinaca sativa", #cam34, 11th July, Pastinaca sativa
    grepl("cam34", cam) & grepl("2024-07-14", date) ~ "Knautia arvensis", #cam34, 14th July, Knautia arvensis
    grepl("cam34", cam) & grepl("2024-07-18", date) ~ "Peucedanum officinale", #cam34, 18th July, Peucedanum officinale
    grepl("cam34", cam) & grepl("2024-07-20", date) ~ "Daucus carota", #cam34, 20th July, Daucus carota
    grepl("cam34", cam) & grepl("2024-07-21", date) ~ "Trifolium pratense", #cam34, 21st July, Trifolium pratense
    grepl("cam34", cam) & grepl("2024-07-24", date) ~ "Trifolium pratense", #cam34, 24th July, Trifolium pratense
    grepl("cam34", cam) & grepl("2024-07-29", date) ~ "Daucus carota", #cam34, 29th July, Daucus carota
    grepl("cam34", cam) & grepl("2024-07-30", date) ~ "Dianthus carthusianorum", #cam34, 30th July, Dianthus carthusianorum
    grepl("cam34", cam) & grepl("2024-08-13", date) ~ "Daucus carota", #cam34, 13th August, Daucus carota
    
    grepl("cam35", cam) & grepl("2024-07-11", date) ~ "Pastinaca sativa", #cam35, 11th July, Pastinaca sativa
    grepl("cam35", cam) & grepl("2024-07-14", date) ~ "Hypericum perforatum", #cam35, 14th July, Hypericum perforatum
    grepl("cam35", cam) & grepl("2024-07-18", date) ~ "Pastinaca sativa", #cam35, 18th July, Pastinaca sativa
    grepl("cam35", cam) & grepl("2024-07-20", date) ~ "Crepis biennis", #cam35, 20th July, Crepis biennis
    grepl("cam35", cam) & grepl("2024-07-21", date) ~ "Pastinaca sativa", #cam35, 21st July, Pastinaca sativa
    grepl("cam35", cam) & grepl("2024-07-24", date) ~ "Hypericum perforatum", #cam35, 24th July, Hypericum perforatum
    grepl("cam35", cam) & grepl("2024-07-29", date) ~ "Pastinaca sativa", #cam35, 29th July, Pastinaca sativa
    grepl("cam35", cam) & grepl("2024-07-30", date) ~ "Daucus carota", #cam35, 30th July, Daucus carota
    grepl("cam35", cam) & grepl("2024-08-13", date) ~ "Trifolium pratense" #cam35, 13th August, Trifolium pratense
  ))
```

```{r save data}
#save the data as a csv file
write_csv(flower_cam, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/bioclip_flower_cams.csv")
```
