---
title: "exploratory_visu"
author: "Alma Steireif"
date: "`r Sys.Date()`"
output: html_document
---

Author: Alma Steireif
Date: `r Sys.Date()`

### Description

The goal of this script is to load and clean the data from the pan traps and transect netting, and to create exploratory visualizations of the data.

For this study, we are only interested in the insect families that are pollinators. We will focus on these families: Coleoptera, Diptera, Hymenoptera and Lepidoptera. For the pan traps, only insects that are larger than 3mm will be considered.


```{r, include=FALSE}
library(tidyverse)
library(viridis)
library(stringr)

library(googlesheets4)
library(readxl)
library(writexl)

library(igraph)
library(ggraph)
library(tidygraph) 

library(iNEXT)
library(fossil) # helps with creating site-by-species-matrix for use with iNEXT 

```



## Pan traps

Steps: 

1. load from excel file
2. remove rows with missing field labels
3. separate the site name into columns for site, transect, color and date
4. format the date column
5. add a column with the type of site (young restored or reference)
6. clean the 'small' column by removing prefixes and splitting counts and insect names
7. pivot the data into wide format
8. sum thrips and small thrips -> during the lab processing of the samples the column "less than 1 mm" was used to count solely thrips that were smaller than 1 mm. This column was then renamed to "thrips" and the counts were added to the "small_thrip" column.
9. removing columns from families that are not taken into consideration in this study. 

```{r load and clean data}
# read one sheet from an excel file
df <-readxl::read_excel("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/data_field/Labels_pantraps_clean.csv.xlsx", sheet = "Labels_pantraps_clean",range = "A1:AR138")

bowltrap_clean <- df %>%
  # Remove rows where 'ID_date' contains '_missing'
  filter(!str_detect(ID_date, "_missing")) %>%
  select(SITE = ID_date, 
         small = notes,
         thrips = less_1mm, #all very small thrips were counted in the "less than 1 mm" column
         apidae:acalyptrate) %>% #select sites' names, notes and all the taxa columns
  #transform full site names from this format "WED_2_Yellow_11.07.2024" into separate columns for site, transect, color and date 
  separate(SITE, 
           into = c("Site", "Transect", "Color", "Date"), 
           sep = "_", 
           remove = FALSE)%>% #keep original column
  mutate(Date = as.Date(Date, format = "%d.%m.%Y"))%>% #format date
  #add column with the type of site
  mutate(Site_type = case_when( #adding site_type column
                Site %in% c("WED", "KOT", "WDG", "BUH") ~ "young_restored", #young restored sites
                Site %in% c("STP", "JEP", "DES", "HLI", "WUP") ~ "reference", #reference sites
                TRUE ~ NA_character_))%>%# Default case, if needed
  relocate(Site_type, .after = Site)  # Move "Site_type" right after "Site"
  
# Clean and process the `small` column
bowltrap_clean <- bowltrap_clean %>%
  mutate(small = str_remove(small, "small: ")) %>%  # Remove "small: " prefix
  mutate(small = str_replace_all(small, "none", "")) %>%  # Replace "none" with an empty string
  separate_rows(small, sep = ", ") %>%  # Split rows into individual entries
  
  # Split counts and insect names
  separate(small, into = c("count", "insect"), sep = " ", extra = "merge", fill = "right") %>%
  
  # Handle invalid `count` values
  mutate(
    count = as.numeric(count),  # Convert `count` to numeric
    insect = ifelse(is.na(count), NA, insect)  # Remove `insect` if `count` is invalid
  ) %>%
  filter(!is.na(count) & !is.na(insect)) %>%  # Remove rows with invalid counts or insects
  
  # Pivot into wide format
  pivot_wider(names_from = insect, values_from = count, values_fill = 0, names_prefix = "small_") %>%

  # Sum thrips and small_thrip
  mutate(small_thrip = coalesce(small_thrip, 0) + coalesce(thrips, 0)) %>%  
  select(-thrips,-small_) %>%  # remove the original thrips column
  
  #sum bombus and bees
  #mutate(symphyta = coalesce(symphyta, 0) + coalesce(tenthredinidae,0)) %>% 
  #mutate(apocrita = coalesce(apocrita, 0) + coalesce(proctotrupidae ,0)) %>%
  #select(-proctotrupidae,-tenthredinidae)%>%
  select(-spider, -cicadellidae, -hemiptera) #remove columns that are not taken into consideration in this study

rm(df) #remove the larger original dataframe to clear up memory
```

```{r save wide format data into csv}}
# remove all columns that start with "small_"
bowltrap_clean_save <- bowltrap_clean %>% 
  select(-starts_with("small_"))%>%
  # Combine "Site" and "Transect" into a new column
  mutate(Site_Tn = paste(Site, Transect, sep = "_")) %>%
  #remove SITE column
  select(-SITE) %>%
  #group rows that have the same Site_Tn value,
  group_by(Site_Tn, Site, Transect,Site_type,Date, ) %>%
  #sum up the counts for each taxa column [,6:30]
  summarise(across(apidae:acalyptrate, sum)) %>%
  #ungroup the data
  ungroup()

#save the data in wide format
write.csv(bowltrap_clean_save, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/bowltrap_clean.csv", row.names = FALSE)

rm(bowltrap_clean_save) #remove the dataframe to clear up memory
```


```{r large insects}
#create subset of data only for larger than 3mm insects
bowl_large <- bowltrap_clean %>%
  select(Site:acalyptrate) %>%
  pivot_longer(cols =  apidae:acalyptrate,  
               names_to = "Taxa", 
               values_to = "Count") 

ggplot(bowl_large, aes(x = Taxa, y = Count, fill = Site)) +
  geom_bar(stat = "identity") +
  labs(title = "Stacked Taxa Counts by Sampling Event",
       x = "Sampling Event",
       y = "Total Count") +
  theme_minimal() +
  scale_fill_viridis(discrete = TRUE, option= "plasma") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 70), n.breaks =8)
```

```{r comparison platform}

#the platform categories are quite different than families, so let's convert these categories to something similar so we can compare

plat_bowl_large <- bowl_large %>%
  mutate(category = case_when(
    Taxa %in% c("cantharidae", "carabidae", "curculionidae", "elateridae","mordellidae", "staphylinidae") ~ "beetle",
    Taxa %in% c("hemiptera") ~ "bug",
    Taxa %in% c("diptera","calliphoridae","cecidomyiidae","tachinidae","calliphoridae","Sepsidae","Ephydridae","Muscidae","Asilidae","Stratiomyidae", "polleniidae","acalyptrate" ) ~ "fly",
    Taxa %in% c("dasypoda","apidae") ~ "bee",
    Taxa %in% c("sarcophagidae") ~ "fly_sarco",
    Taxa %in% c("symphyta","apocrita") ~ "wasp",
    Taxa %in% c("Empididae") ~ "fly_empi",
    TRUE ~ Taxa
  ))

ggplot(plat_bowl_large, aes(x = reorder(category, -Count), y = Count, fill = Site)) +
  geom_bar(stat = "identity",na.rm=TRUE ) +
  labs(title = "Pan traps: Stacked Taxa Counts per Site",
       x = "Taxa",
       y = "Total Count") +
  theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 70))
```

```{r small insects}
bowl_small <- bowltrap_clean %>%
  select(Site:Date, small_thrip:small_staphylinidae) %>%
  
  mutate(small_diptera = coalesce(small_fly, 0) + coalesce(small_cecidomyiidae, 0)) %>%  
  mutate(small_coleoptera = coalesce(small_beetle, 0) + coalesce(small_staphylinidae, 0)) %>%
  mutate(small_hemiptera= coalesce(small_aphid, 0) + coalesce(small_hemiptera, 0)) %>%
  
  select(-small_fly, -small_cecidomyiidae, -small_aphid, -small_staphylinidae, -small_beetle) %>%
  
  pivot_longer(cols =  small_thrip:last_col(),  
               names_to = "Taxa",
               values_to = "Count")
#plots
ggplot(bowl_small, aes(x = Site, y = Count, fill = Taxa)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Taxa Counts by Sampling Event",
       x = "Taxa",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(bowl_small, aes(x = Site, y = Count, fill = Taxa)) +
  geom_bar(stat = "identity") +
  labs(title = "Stacked taxa Counts by Sampling Event",
       x = "Sampling Event",
       y = "Total Count") +
  theme_minimal() +
  scale_fill_viridis(discrete = TRUE, option= "plasma") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### iNext - rarefaction curves
- iNEXT focuses on three measures of Hill numbers of order q: 
    + species richness (q = 0), 
    + Shannon diversity (q = 1, the exponential of Shannon entropy) 
    + Simpson diversity (q = 2, the inverse of Simpson concentration)    
 
[Source](https://cran.r-project.org/web/packages/iNEXT/vignettes/Introduction.html) 

```{r rarefaction curves for pan traps}
bowl_large2 <- as.data.frame(bowl_large) #make 'dat' into a data frame that can be read by create.matrix
bowl_large2 <-subset(bowl_large2, bowl_large2[6]!=0) #remove rows with empty values in the 'family' column

#rarefaction curves for site_type ---------------------------------------------
matrix_bowl_type <- create.matrix(x = bowl_large2,
                                tax.name = 'Taxa',
                                locality = 'Site_type',
                                abund = TRUE,
                                abund.col = 'Count')
##Check your work
identical(sum(matrix_bowl_type),sum(bowl_large2$Count))
#both have 105 interactions

# Convert to data.frame for use with iNEXT
matrix_bowl_type <- as.data.frame(matrix_bowl_type)

#run iNEXT to get the diversity estimates per SITE_TYPE
set.seed(2022);iNEXT_output_insects <- iNEXT(matrix_bowl_type,  q = c(0,1,2), datatype = 'abundance') #set seed for reproducibility, iNEXT include all three indices of diversity

# This are specifically the diversity estimates with their confidence intervals:
# iNEXT_output_insects$AsyEst 

#Plot the results and save the plot
(rarefaction_bowl_sitetype <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage',facet.var="Order.q")+
  ggtitle("Pan traps: rarefaction curves per site type")+
  labs(y= "Family diversity"))
ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_bowl_sitetype.png")

rm(matrix_bowl_type, rarefaction_bowl_sitetype) #remove the matrix and plot to clear up memory

#rarefaction curves for site ---------------------------------------------
matrix_bowl_site <- create.matrix(x = bowl_large2,
                                tax.name = 'Taxa',
                                locality = 'Site',
                                abund = TRUE,
                                abund.col = 'Count')
##Check your work
identical(sum(matrix_bowl_site),sum(bowl_large2$Count))
#both have 105 interactions

# Convert to data.frame for use with iNEXT
matrix_bowl_site <- as.data.frame(matrix_bowl_site)

#run iNEXT to get the diversity estimates per SITE
set.seed(2022);iNEXT_output_insects <- iNEXT(matrix_bowl_site,  q = c(0,1,2), datatype = 'abundance')  #set seed for reproducibility, iNEXT include all three indices of diversity

# This are specifically the diversity estimates with their confidence intervals:
# iNEXT_output_insects$AsyEst

# Plot the results and save the plot
(rarefaction_bowl_site <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage', facet.var = "Order.q") +
  scale_shape_manual(values = c(0, 1, 2, 3, 4, 5, 6, 7, 8))+ # Assign custom shapes
  ggtitle("Pan traps: rarefaction curves per site")+
  labs(y= "Family diversity"))
ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_bowl_site.png")

rm(matrix_bowl_site, rarefaction_bowl_site) #remove the matrix and plot to clear up memory

#rarefaction curves for transect ---------------------------------------------  
matrix_bowl_transect <- create.matrix(x = bowl_large2,
                                tax.name = 'Taxa',
                                locality = 'Transect',
                                abund = TRUE,
                                abund.col = 'Count')

##Check your work
identical(sum(matrix_bowl_transect),sum(bowl_large2$Count)) #both have 105 interactions

# Convert to data.frame for use with iNEXT
matrix_bowl_transect <- as.data.frame(matrix_bowl_transect)

#run iNEXT to get the diversity estimates per transect number
set.seed(2022);iNEXT_output_insects <- iNEXT(matrix_bowl_transect,  q = c(0,1,2), datatype = 'abundance') #set seed for reproducibility, iNEXT include all three indices of diversity

# This are specifically the diversity estimates with their confidence intervals:
# iNEXT_output_insects$AsyEst

# plot the results and save the plot
(rarefaction_bowl_transect <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage',facet.var="Order.q") +
  ggtitle("Pan traps: rarefaction curves per transects")+
  labs(y= "Family diversity"))
ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_bowl_transect.png")

rm(matrix_bowl_transect, rarefaction_bowl_transect) #remove the matrix and plot to clear up memory

#rarefaction curves for bowl color ---------------------------------------------
matrix_bowl_color <- create.matrix(x = bowl_large2,
                                tax.name = 'Taxa',
                                locality = 'Color',
                                abund = TRUE,
                                abund.col = 'Count')

##Check your work
identical(sum(matrix_bowl_color),sum(bowl_large2$Count)) #both have 105 interactions

# Convert to data.frame for use with iNEXT
matrix_bowl_color <- as.data.frame(matrix_bowl_color)

set.seed(2022);iNEXT_output_insects <- iNEXT(matrix_bowl_color,  q = c(0,1,2), datatype = 'abundance') 
#set seed for reproducibility, iNEXT include all three indices of diversity
# iNEXT_output_insects$AsyEst # AsyEst gives the diversity estimates with their confidence intervals

# Plot the results and save the plot
(rarefaction_bowl_color <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage',facet.var="Order.q")+
  ggtitle("Pan traps: rarefaction curves per bowl color")+
  labs(y= "Family diversity"))
ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_bowl_color.png")

rm(matrix_bowl_color, rarefaction_bowl_color) #remove the matrix and plot to clear up memory
```

## Transect netting

```{r load transect data,include=FALSE}
#load data
 #Read google sheets data into R
net_data_load <-read_sheet("https://docs.google.com/spreadsheets/d/1YDOoUobRU6A36um0-iut6kY77X0jEMYXtFWR3-Zankc/edit?usp=sharing", sheet=2) #sheet 2 has the interaction data
```

```{r clean transect data}
net_data <- net_data_load %>%
  select(date=date, site = Site, transect = Transect, final_ID = final_ID, plant_species=plant_species, field_ID=pollinator_field_ID, observed =observed, collected=collected, family=Family, genus= genus, species=final_species) #select columns of interest


net_data <- net_data %>%
        mutate(site_type = case_when( #adding site_type column
                site %in% c("WED", "KOT", "WDG", "BUH") ~ "young_restored", #young restored sites
                site %in% c("STP", "JEP", "DES", "HLI", "WUP") ~ "reference", #reference sites
                TRUE ~ NA_character_)) %>%# Default case, if needed
        # new column with observed + collected interactions = total interactions
        mutate(total_interaction = coalesce(observed, 0) + coalesce(collected, 0))%>% 
        # Standardize the `Family` column to unify textual variations
        mutate(family = case_when(
               family %in% c("Muscidae c.f.", "Muscidae (c.f)", "Muscidae") ~ "Muscidae",  
               family %in% c("Tachinidae", "Tachinidae (c.f)") ~ "Tachinidae",
               family %in% c("Calliphoridae", "Calliphoridae c.f.") ~ "Calliphoridae",
               family %in% c("Mellitidae", "Melittidae") ~ "Melittidae",
               family %in% c("Fanniidae", "Faniidae") ~ "Fanniidae",
               family %in% c("Sarcophagidae c.f.", "Sarcophagidae") ~ "Sarcophagidae",
               family %in% c("Tenthrenidae", "Tenthredinidae") ~ "Tenthredinidae",
               TRUE ~ family)) %>% # Keep other families unchanged
        filter(!is.na(family)) %>% #remove rows with empty values in the 'family' column
        filter(family!="other") %>% #remove rows with the value "other" in the 'family' column
        filter(site!="") %>%#remove rows with empty values in the 'site' column
  
  # Standardize the genus column to unify textual variations
        mutate(genus= case_when(
               genus %in% c("aff. Syrphus", "Syrphus") ~ "Syrphus",
               genus %in% c("aff. Eupeodes", "Eupeodes") ~ "Eupeodes",
               TRUE ~ genus)) %>% # Keep other lowest_taxa unchanged
  
        # Create a new column that defines the lowest taxonomic level available composed of all taxonomic levels available
       mutate(
         species = gsub(" ", "_", trimws(species)),  # Replace spaces in species names
         genus = trimws(genus),
         family = trimws(family),
         lowest_taxa = case_when( species != "" ~ paste(family, species, sep = "_"),  # Use family_species if species is available
                                  genus != "" ~ paste(family, genus, sep = "_"),      # Use family_genus if only genus is available
                                  TRUE ~ family)) %>%                                # Keep just family if no genus or species
        
        #remove  rows that have no info in plant_species 
        filter(!(plant_species=="no info"))
```

```{r save cleaned data into csv}

### FAMILY LEVEL

#select columns of interest and create new column that contains the site_transect_plant_species
net_data_save_fam <- net_data %>%
  select(site,  site_type, transect, date, plant_species, total_interaction, family)%>%
  #create new column that contains the site_transect_plant_species
  mutate(Site_Tn_plantsp = paste(site, transect, plant_species, sep = "_")) %>%
  #add _ where spaces are in the Site_Tn_plantsp column
  mutate(Site_Tn_plantsp = gsub(" ", "_", Site_Tn_plantsp))%>%
  
  ## PIVOT STEPS
  #remove empty rows in the 'family' column
  filter(!is.na(family)) %>%
  
  # pivot wider - taking family for headers and total_interaction for values, and site_transect_plant_species as rows
  pivot_wider(
    names_from = family, 
    values_from = total_interaction, 
    values_fill = list(total_interaction = 0),  # Fill missing values with 0
    values_fn = list(total_interaction = sum)  # Sum values when there are duplicates
  ) 

#save the data in wide format
write.csv(net_data_save_fam, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/net_data_family.csv", row.names = FALSE)


### TAXA LEVEL

net_data_save_low <- net_data %>%
  select(date, site, transect, site_type, plant_species, total_interaction, lowest_taxa)%>%
  #create new column that contains the site_transect_plant_species
  mutate(Site_Tn_plantsp = paste(site, transect, plant_species, sep = "_")) %>%
  #add _ where spaces are in the Site_Tn_plantsp column
  mutate(Site_Tn_plantsp = gsub(" ", "_", Site_Tn_plantsp))%>%
  
  ## PIVOT STEPS
  # pivot wider - taking lowest_taxa for headers and total_interaction for values, and site_transect_plant_species as rows
  pivot_wider(
    names_from = lowest_taxa, 
    values_from = total_interaction, 
    values_fill = list(total_interaction = 0),  # Fill missing values with 0
    values_fn = list(total_interaction = sum)  # Sum values when there are duplicates
  )

# save the data in wide format
write.csv(net_data_save_low, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/net_data_taxa.csv", row.names = FALSE)

rm(net_data_save_fam, net_data_save_low) #remove the data frames to clear up memory
```




```{r interaction counts and plot}
# ~ lowest taxonomic level available ----------------------------------------------
interaction_counts_site_low <- net_data %>%
  # Group by site and lowest taxonomic level, then count interactions
  group_by(site, lowest_taxa) %>%
  summarize(interaction_count = n(), .groups = 'drop') # Summarize the data

ggplot(interaction_counts_site_low, aes(x = site, y = interaction_count, fill = lowest_taxa)) +
  geom_bar(stat = "identity",colour="white") +
  labs(title = "Stacked lowest_taxa Counts by Sampling Event",
       x = "Sampling Event",
       y = "Total Count") +
  theme_minimal() +
  scale_fill_viridis(discrete = TRUE,guide = guide_legend(ncol = 3)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ 
  #theme(legend.position = "none")+
  theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title = element_text(size=12), #change legend title font size
        legend.text = element_text(size=8)) #change legend text font size
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Interaction counts by site ----------------------------------------------
# Summarize to get the count of species per site
interaction_counts_site <- net_data %>%
  # Group by SITE and standardized family name (fam), then count interactions
  group_by(site, family) %>%
  summarize(interaction_count = n(), .groups = 'drop') #summarize the data
  
ggplot(interaction_counts_site, aes(x = site, y = interaction_count, fill = family)) +
  geom_bar(stat = "identity") +
  labs(title = "Stacked Species Counts by Sampling Event",
       x = "Sampling Event",
       y = "Total Count") +
  theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Interaction counts by plant species ---------------------------------------------- 
# Summarize to get the count of interaction between insect and plant species
interaction_counts_plants <- net_data %>%
  # Group by SITE and standardized family name (fam), then count interactions
  group_by(plant_species, family) %>%
  summarize(interaction_count = n(), .groups = 'drop')
  
# plot the data according to plant species
ggplot(interaction_counts_plants, aes(x = reorder(plant_species, interaction_count), y = interaction_count, fill = family)) +
  geom_bar(stat = "identity") +
  labs(title = "Stacked Species Counts by flower species",
       x = "Site",
       y = "Total Count") +
  theme_minimal() +
  scale_fill_viridis(discrete = TRUE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  coord_flip()
rm(interaction_counts_plants)
```

```{r transect plot}
# Create a frequency table for pol_ID
fam_counts <- table(interaction_counts_site$family)

# Convert the table to a dataframe
fam_counts_df <- as.data.frame(table(interaction_counts_site$family))
colnames(fam_counts_df) <- c("fam_ID", "frequency")

# Create the bar plot
ggplot(fam_counts_df, aes(x = fam_ID, y = frequency, width=.7)) +
  geom_col(fill= "skyblue") + # Create the bar plot
  geom_text(aes(label = frequency), vjust = -0.5, size = 4) + # Add labels above bars
  labs(
    title = "Counts per Family",
    x = "Pollinator ID",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

sum(fam_counts_df$Frequency);length(fam_counts_df$fam_ID);levels(as.factor(interaction_counts_site$family))
rm(fam_counts, fam_counts_df,interaction_counts_site)
```

### iNext - rarefaction curves
```{r rarefaction curves for transect netting}
# rarefaction curves for site_type ---------------------------------------------
net_data <- as.data.frame(net_data) #make 'dat' into a data frame that can be read by create.matrix

net_data<-subset(net_data, net_data[8]!="") #remove rows with empty values in the 'family' column

matrix_netting_type <- create.matrix(x = net_data,
                                tax.name = 'family',
                                locality = 'site_type',
                                abund = TRUE,
                                abund.col = 'total_interaction')
##Check your work
identical(sum(matrix_netting_type),sum(net_data$total_interaction))
#both have 2427 interactions

# Convert to data.frame for use with iNEXT
matrix_netting_type <- as.data.frame(matrix_netting_type)

# Run iNEXT to get the diversity estimates per SITE_TYPE
set.seed(2022);iNEXT_output_insects <- iNEXT(matrix_netting_type,  q = c(0,1,2), datatype = 'abundance')#set seed for reproducibility, iNEXT include all three indices of diversity

# this are specifically the diversity estimates with their confidence intervals:
# iNEXT_output_insects$AsyEst 

# Plot the results and save the plot
(rarefaction_netting_sitetype <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage',facet.var="Order.q")+
  ggtitle("Netting: rarefaction curves per site type")+
  labs(y= "Family diversity"))
ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_netting_sitetype.png")

rm(matrix_netting_type, rarefaction_netting_sitetype)

# rarefaction curves for sites ----------------------------------------------
matrix_netting_site <- create.matrix(x = net_data,
                                tax.name = 'family',
                                locality = 'site',
                                abund = TRUE,
                                abund.col = 'total_interaction')
##Check your work
identical(sum(matrix_netting_site),sum(net_data$total_interaction))
#both have 2427 interactions

# Convert to data.frame for use with iNEXT
matrix_netting_site <- as.data.frame(matrix_netting_site)

# Run iNEXT to get the diversity estimates per SITE 
set.seed(2022); iNEXT_output_insects <- iNEXT(matrix_netting_site,  q = c(0,1,2), datatype = 'abundance') #set seed for reproducibility, iNEXT include all three indices of diversity

# this are specifically the diversity estimates with their confidence intervals:
# iNEXT_output_insects$AsyEst 

# Plot the results and save the plot
(rarefaction_netting_site <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage',facet.var="Order.q")+
  ggtitle("Netting: rarefaction curves per site")+
  labs(y= "Family diversity"))
ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_netting_site.png")

rm(matrix_netting_site, rarefaction_netting_site) #remove the matrix and plot to clear up memory

# rarefaction curves for transect number ---------------------------------------------
matrix_netting_transect <- create.matrix(x = net_data,
                                tax.name = 'family',
                                locality = 'transect',
                                abund = TRUE,
                                abund.col = 'total_interaction')
##Check your work
identical(sum(matrix_netting_transect),sum(net_data$total_interaction))
#both have 2427 interactions

# Convert to data.frame for use with iNEXT
matrix_netting_transect <- as.data.frame(matrix_netting_transect)

#run iNEXT to get the diversity estimates per transect number
set.seed(2022); iNEXT_output_insects <- iNEXT(matrix_netting_transect,  q = c(0,1,2), datatype = 'abundance') #set seed for reproducibility, iNEXT include all three indices of diversity

# This are specifically the diversity estimates with their confidence intervals:
# iNEXT_output_insects$AsyEst 

# plot the results and save the plot
(rarefaction_netting_transect <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage',facet.var="Order.q")+
  ggtitle("Netting: rarefaction curves per transects")+
  labs(y= "Family diversity"))
ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_netting_transect.png")

rm(matrix_netting_transect, rarefaction_netting_transect) #remove the matrix and plot to clear up memory 
```

### Rarefaction curves for lowest taxa

- since not all insects were identified to the same level, but we still want to get as much diversity information as we can, we will use the lowest taxonomic level available for each insect.

```{r rarefaction curves for transect netting lowest taxa (lt)}
# rarefaction curves for site_type_lt ---------------------------------------------
net_data <- as.data.frame(net_data) #make 'dat' into a data frame that can be read by create.matrix

net_data<-subset(net_data, net_data[13]!="") #remove rows with empty values in the 'family' column

matrix_netting_type <- create.matrix(x = net_data,
                                tax.name = 'lowest_taxa',
                                locality = 'site_type',
                                abund = TRUE,
                                abund.col = 'total_interaction')
##Check your work
identical(sum(matrix_netting_type),sum(net_data$total_interaction))
#both have 2427 interactions

# Convert to data.frame for use with iNEXT
matrix_netting_type <- as.data.frame(matrix_netting_type)

# Run iNEXT to get the diversity estimates per SITE_TYPE
set.seed(2022);iNEXT_output_insects <- iNEXT(matrix_netting_type,  q = c(0,1,2), datatype = 'abundance')#set seed for reproducibility, iNEXT include all three indices of diversity

# this are specifically the diversity estimates with their confidence intervals:
# iNEXT_output_insects$AsyEst 

# Plot result and save the plot
(rarefaction_netting_sitetype_lt <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage',facet.var="Order.q")+
  ggtitle("Netting: rarefaction curves per site type")+
  labs(y= "Lowest taxa diversity"))

  ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_netting_sitetype_lt.png")
  
# rarefaction curves for sites ----------------------------------------------
matrix_netting_site_lt <- create.matrix(x = net_data,
                                tax.name = 'lowest_taxa',
                                locality = 'site',
                                abund = TRUE,
                                abund.col = 'total_interaction')
##Check your work
identical(sum(matrix_netting_site_lt),sum(net_data$total_interaction))
#both have 2427 interactions

# Convert to data.frame for use with iNEXT
matrix_netting_site_lt <- as.data.frame(matrix_netting_site_lt)

# Run iNEXT to get the diversity estimates per SITE 
set.seed(2022); iNEXT_output_insects <- iNEXT(matrix_netting_site_lt,  q = c(0,1,2), datatype = 'abundance') #set seed for reproducibility, iNEXT include all three indices of diversity

# this are specifically the diversity estimates with their confidence intervals:
# iNEXT_output_insects$AsyEst 

# Plot the results and save the plot
(rarefaction_netting_site_lt <- ggiNEXT(iNEXT_output_insects, type = 1, color.var = 'Assemblage',facet.var="Order.q")+
  ggtitle("Netting: rarefaction curves per site")+
  labs(y= "Lowest taxa diversity"))
ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_netting_site_lt.png")

rm(matrix_netting_site, rarefaction_netting_site) #remove the matrix and plot to clear up memory



rm(matrix_netting_type, rarefaction_netting_sitetype_lt)
```



```{r}
rm(iNEXT_output_insects,bowl_large2)
```

```{r saving dataframes}
# net_data ---------------------------------------------
# save a RData file
save(net_data, file = paste0("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/", current_date, "_netdata.RData"))
#Save as excel file
#write_xlsx(net_data, path = paste0("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/", current_date, "_netdata.xlsx"))

# bowl_large ---------------------------------------------
# save a RData file
#save(bowl_large, file = paste0("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/", current_date, "_bowl_large.RData"))
#Save as excel file
#write_xlsx(bowl_large, path = paste0("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/", current_date, "_bowl_large.xlsx"))
```

