---
title: "camtrap_analysis_Alma"
author: "Alma Steireif"
date: "`r Sys.Date()`"
output: html_document
---

 
Original Author:  Maximilian Sittinger (https://orcid.org/0000-0002-4096-8556)
Modifications: Alma Steireif

This R script does the following:
 - process and plot data from 8 Insect Detect camera traps, which were deployed above platforms in 9 different meadows in Sachsen-Anhalt, Germany from July to August 2024
 
 
## 1 Import required packages

```{r setup, include=FALSE}
# Load required packages
library(tidyverse)
library(patchwork)
library(viridis)

sessionInfo()
```
 
Script was written and plots were created with:
 - R version 4.4.0 (2024-04-24 ucrt)
 - viridis_0.6.5, viridisLite_0.4.2,mpatchwork_1.3.0, lubridate_1.9.3, forcats_1.0.0, stringr_1.5.1, dplyr_1.1.4, purrr_1.0.2, readr_2.1.5, tidyr_1.3.1, tibble_3.2.1, ggplot2_3.5.1, tidyverse_2.0.0  

# Platform cameras

## 2 Import platform data

Import camera trap metadata with post-processed classification results
 - tracking IDs with less than 3 images and 85 tracking IDs with more than 1,800 images were removed during post-processing (default settings: "-min_tracks 3" and “-max_tracks 1800”)
 
```{r load platform data}
camtrap36 <- read_csv("C:/Users/Almas/YOLOv5/yolov5-master/runs/predict-cls/seppi-cam36/results/2024-10-15_15-05-49_metadata_processed/seppi-cam36_metadata_classified_top1_final.csv", col_names = TRUE)
camtrap37 <- read_csv("C:/Users/Almas/YOLOv5/yolov5-master/runs/predict-cls/seppi-cam37/results/2024-10-15_15-09-59_metadata_processed/seppi-cam37_metadata_classified_top1_final.csv", col_names = TRUE)
camtrap38 <- read_csv("C:/Users/Almas/YOLOv5/yolov5-master/runs/predict-cls/seppi-cam38/results/2024-10-15_15-01-34_metadata_processed/seppi-cam38_metadata_classified_top1_final.csv", col_names = TRUE)

# Combine data from all camera traps
platform_cams <- bind_rows(
  camtrap36,
  camtrap37,
  camtrap38
)

rm(camtrap36, camtrap37, camtrap38) # Remove individual cameras dataframes
```

```{r site assignment platform}
#associate experiment date with site
platform_cams <- platform_cams %>%
  mutate(location = case_when(
    grepl("2024-07-11", start_time) ~ "WED",  # Match the 11th July to Wedringen
    grepl("2024-07-14", start_time) ~ "STP",  # Match the 14th July to Steutz
    grepl("2024-07-18", start_time) ~ "JEP",  # Match the 18th July to Jerichow
    grepl("2024-07-20", start_time) ~ "KOT",  # Match the 20th July to Köthen
    grepl("2024-07-21", start_time) ~ "DES",  # Match the 21nd July to Dessau
    grepl("2024-07-24", start_time) ~ "HLI",  # Match the 24th July to Pölsfeld
    grepl("2024-07-29", start_time) ~ "WUP",  # Match the 29th July to Wulfen
    grepl("2024-07-30", start_time) ~ "WDG",  # Match the 30th July to Wedringen
    grepl("2024-08-13", start_time) ~ "BUH",  # Match the 13th August to Bühne
    TRUE ~ "UNKNOWN"  # Default value
  ))
```

```{r transect assignment platform}
platform_cams <- platform_cams %>%
  mutate(transect = case_when(
    grepl("seppi-cam36", cam_ID) & grepl("WED", location) ~ "3",  #cam36, WED, T3
    grepl("seppi-cam36", cam_ID) & grepl("STP", location) ~ "1",  #cam36, STP, T4
    grepl("seppi-cam36", cam_ID) & grepl("JEP", location) ~ "2",  #cam36, JEP, T2
    grepl("seppi-cam36", cam_ID) & grepl("KOT", location) ~ "5",  #cam36, KOT, T5
    grepl("seppi-cam36", cam_ID) & grepl("DES", location) ~ "1",  #cam36, DES, T1
    grepl("seppi-cam36", cam_ID) & grepl("HLI", location) ~ "1",  #cam36, HLI, T1
    grepl("seppi-cam36", cam_ID) & grepl("WUP", location) ~ "3",  #cam36, WUP, T3
    grepl("seppi-cam36", cam_ID) & grepl("WDG", location) ~ "5",  #cam36, WDG, T5
    grepl("seppi-cam36", cam_ID) & grepl("BUH", location) ~ "1",  #cam36, BUH, T1
    
    grepl("seppi-cam37", cam_ID) & grepl("WED", location) ~ "5",  #cam37, WED, T5
    grepl("seppi-cam37", cam_ID) & grepl("STP", location) ~ "5",  #cam37, STP, T5
    grepl("seppi-cam37", cam_ID) & grepl("JEP", location) ~ "5",  #cam37, JEP, T5
    grepl("seppi-cam37", cam_ID) & grepl("KOT", location) ~ "1",  #cam37, KOT, T1
    grepl("seppi-cam37", cam_ID) & grepl("DES", location) ~ "3",  #cam37, DES, T3
    grepl("seppi-cam37", cam_ID) & grepl("HLI", location) ~ "4",  #cam37, HLI, T4
    grepl("seppi-cam37", cam_ID) & grepl("WUP", location) ~ "5",  #cam37, WUP, T5
    grepl("seppi-cam37", cam_ID) & grepl("WDG", location) ~ "1",  #cam37, WDG, T1
    grepl("seppi-cam37", cam_ID) & grepl("BUH", location) ~ "3",  #cam37, BUH, T3
    
    grepl("seppi-cam38", cam_ID) & grepl("WED", location) ~ "2",  #cam38, WED, T2
    grepl("seppi-cam38", cam_ID) & grepl("STP", location) ~ "3",  #cam38, STP, T3
    grepl("seppi-cam38", cam_ID) & grepl("JEP", location) ~ "1",  #cam38, JEP, T1
    grepl("seppi-cam38", cam_ID) & grepl("KOT", location) ~ "3",  #cam38, KOT, T3
    grepl("seppi-cam38", cam_ID) & grepl("DES", location) ~ "5",  #cam38, DES, T5
    grepl("seppi-cam38", cam_ID) & grepl("HLI", location) ~ "5",  #cam38, HLI, T5
    grepl("seppi-cam38", cam_ID) & grepl("WUP", location) ~ "1",  #cam38, WUP, T1
    grepl("seppi-cam38", cam_ID) & grepl("WDG", location) ~ "3",  #cam38, WDG, T3
    grepl("seppi-cam38", cam_ID) & grepl("BUH", location) ~ "5",  #cam38, BUH, T5
    
    TRUE ~ "UNKNOWN"  # Default value
  ))
```
 

 
 
```{r load recording logs platform}
# Import record logs which contain start and end time of each recording, as well as the total minutes recorded per session, the number of crops and IDs 

camtrap36_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam36/record_log.csv", col_names=T) |> mutate(
    # Extract date and time components from rec_start and rec_end
    record_start_date = as_date(rec_start), # Convert to Date
    record_start_time = format(rec_start, "%H:%M"), # Extract Time as HH:MM
    record_end_time = format(rec_end, "%H:%M")) # Extract Time as HH:MM


camtrap37_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam37/record_log.csv", col_names=T) |> mutate(
    # Extract date and time components from rec_start and rec_end
    record_start_date = as_date(rec_start), # Convert to Date
    record_start_time = format(rec_start, "%H:%M"), # Extract Time as HH:MM
    record_end_time = format(rec_end, "%H:%M") # Extract Time as HH:MM
  )

camtrap38_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam38/record_log.csv", col_names=T) |> mutate(
    # Extract date and time components from rec_start and rec_end
    record_start_date = as_date(rec_start), # Convert to Date
    record_start_time = format(rec_start, "%H:%M"), # Extract Time as HH:MM
    record_end_time = format(rec_end, "%H:%M") # Extract Time as HH:MM
  )

#camtrap36_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam36/record_log.csv", col_names= T)
#camtrap37_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam37/record_log.csv", col_names= T)
#camtrap38_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam38/record_log.csv", col_names= T)
```


```{r combine recording logs platform}
#bind the three camera recording logs together and remove the individual dataframes
camtrap_logs_rec <- bind_rows(camtrap36_logs_rec, camtrap37_logs_rec, camtrap38_logs_rec)
#rm(camtrap36_logs_rec, camtrap37_logs_rec, camtrap38_logs_rec)
```
 
# 3 Platform data summary

```{r data summary platform}

# Total recording time of all camera traps
sum(camtrap_logs_rec$rec_time_min) / 60
# ~146.204h

# Total recording time per camera trap
camtrap_logs_rec |> 
  group_by(cam_ID) |> 
  summarise(record_time_total_h = sum(rec_time_min) / 60)
#seppi-cam36	40.35h			
#seppi-cam37	52.79h			
#seppi-cam38	53.05h	

# Total number of captured images
sum(platform_cams$track_ID_imgs)
# 38450 images


# Total number of all unique tracking IDs
nrow(platform_cams)
# 2096  tracking IDs total

# Total number of all unique tracking IDs of insect classes
platform_cams |> 
  filter(str_detect(top1, "none", negate = TRUE)) |> # how many insect pictures are left when excluding the none_something classes
  nrow()
# 1180 tracking IDs of insect classes

# Total number of all unique tracking IDs of non-insect classes
platform_cams |> 
  filter(str_detect(top1, "none")) |> # how many none_something pictures are there
  nrow()
# 916 tracking IDs of non-insect classes (background, bird, dirt, shadow)

  
# Total number of captured images, for which the top1 class with the highest weighted probability was hoverfly
platform_cams |> 
  filter(str_detect(top1, "hfly")) |>
  summarise(track_ID_imgs_hfly = sum(track_ID_imgs))
# 1452 images of hoverfly classes

# Total number of all unique tracking IDs that were classified as hoverfly
platform_cams |> 
  filter(str_detect(top1, "hfly")) |>
  nrow()
# 77 tracking IDs of hoverfly classes

# Total number of all unique tracking IDs per hoverfly class
platform_cams |> 
  filter(str_detect(top1, "hfly")) |>
  count(top1)
#hfly_episyr	22 tracking IDs
#hfly_eristal	4 tracking IDs
#hfly_eupeo	5 tracking IDs
#hfly_myathr	4	tracking IDs
#hfly_sphaero	2 tracking IDs
#hfly_syrphus	40 tracking IDs

table(platform_cams$top1)
```
 
## 4 Set ggplot2 theme 
```{r set ggplot2 theme}
# Save current default theme
theme_default <- theme_get()

# Set new theme
theme_new <- theme_classic(base_size = 12) +
  theme(axis.line = element_blank(),
        axis.text = element_text(colour = "black"),
        axis.ticks = element_line(linewidth = 0.4, colour = "black"),
        legend.key.size = unit(0.5, "cm"),
        legend.margin = margin(t = 0),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 9),
        panel.border = element_rect(linewidth = 0.4, colour = "black", fill = NA),
        panel.grid.major.y = element_line(colour = "grey90", linewidth = 0.2),
        plot.margin = margin(2, 2, 2, 2, "pt"),
        plot.title = element_text(size = 10))
theme_set(theme_new)

# Optional: Set time locale to English to plot month names in correct format
Sys.getlocale()
#Sys.setlocale("LC_TIME", "en_GB.utf8")
```


## 5 Platform plotting

```{r recording duration per session platform}
#visualize the recording duration per session
ggplot(camtrap_logs_rec, aes(x = cam_ID, y = rec_time_min/60, #divide by 60 to get hours
                             fill=rec_time_min/60)) + #fill the bars with the total hours recorded
  geom_bar(stat= "identity") +
  geom_col(colour = "grey")+ #add grey outline to the stacked sections
  labs(title = "Recording duration per session", # Session not day bc some recordings are split within one day
       x = "Camera",
       y = "Total hours recorded") +
  labs(fill = "Hours recorded") +
  theme_minimal()+
  scale_fill_viridis(direction = -1) #reverse the viridis color scale
```


```{r unique IDs per class platform}
# Generate color palette to plot the same color for each month respectively,
# even if the number of months the camera traps were active differs
viridis_9 <- viridis_pal()(9)
#scales::show_col(viridis_5)

# barplot_total function to plot the total number of unique tracking IDs per class
barplot_total <- function(df, class_exclude, title) {
  df |> 
    filter(str_detect(top1, {{ class_exclude }}, negate = TRUE)) |>
    add_count(top1) |> 
    ggplot(aes(reorder(top1, -n))) +
    geom_bar(aes(fill = location), color = "black", linewidth = 0.2) +
    scale_fill_manual(name = "Location", values = viridis_9) +
    labs(title = {{ title }}, x = "Predicted class", y = "Number of unique tracking IDs") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3),
          legend.position = c(0.9, 0.7))
}

(ct36 <- platform_cams |>
  filter(str_detect(cam_ID, "cam36")) |>
  barplot_total(class_exclude = "none", title = "Camtrap 36") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 250), n.breaks = 16))

(ct37 <- platform_cams |>
  filter(str_detect(cam_ID, "cam37")) |>
  barplot_total(class_exclude = "none", title = "Camtrap 37") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 250), n.breaks = 16))

(ct38 <- platform_cams |>
  filter(str_detect(cam_ID, "cam38")) |>
  barplot_total(class_exclude = "none", title = "Camtrap 38") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 70), n.breaks = 16))

(ct_all <- platform_cams |>
  barplot_total(class_exclude = "none", title = "All camera traps") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 550), n.breaks = 16))
  
  
# Combine plots
(ct36 + labs(x = "")) +
(ct37 + labs(x = "", y = "")) +
(ct38 + labs(x = "", y = "")) +
(ct_all + labs(y = "")) +
  plot_annotation(tag_levels = "A") & 
  theme(plot.tag = element_text(size = 11, face = "bold"))

# Plot all classes, including "none_bg", "none_bird", "none_dirt", "none_shadow"
platform_cams |> 
  barplot_total(class_exclude = "0", title = "Camtraps merged") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 550), n.breaks = 14)

```


### Time difference between tracking IDs 
Calculate minimum difference between end_time and start_time of the 5 previous tracking IDs of the same predicted class, grouped by camtrap, date, recording interval.

```{r time difference between tracked IDs platform}

timediff <- platform_cams |> 
  filter(str_detect(top1, "none", negate = TRUE)) |>
  group_by(cam_ID, date, hour(start_time), top1) |> 
  mutate(time_diff_prevID_1 = start_time - lag(end_time, 1)) |> 
  mutate(time_diff_prevID_2 = start_time - lag(end_time, 2)) |> 
  mutate(time_diff_prevID_3 = start_time - lag(end_time, 3)) |> 
  mutate(time_diff_prevID_4 = start_time - lag(end_time, 4)) |> 
  mutate(time_diff_prevID_5 = start_time - lag(end_time, 5)) |> 
  mutate(across(time_diff_prevID_1:time_diff_prevID_5, ~replace(., . < 0, NA))) |> 
  mutate(time_diff_prevID = pmin(time_diff_prevID_1, time_diff_prevID_2, time_diff_prevID_3,
                                 time_diff_prevID_4, time_diff_prevID_5, na.rm = TRUE)) |>
  filter(time_diff_prevID < 30) # keep only time differences < 30 s

# Number of tracking IDs with time difference of < 5 s to previous ID of the same top1 class
timediff |> 
  filter(time_diff_prevID < 5) |> 
  nrow()
# 204 tracking IDs

# Plot time difference to previous tracking ID of the same predicted class
timediff |> 
  ggplot(aes(as.numeric(time_diff_prevID))) +
  geom_histogram(aes(fill = as.factor(cam_ID)), binwidth = 1, color = "black", linewidth = 0.2) +
  scale_fill_viridis_d(name = "Camtrap") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100), n.breaks = 11) +
  scale_x_continuous(expand = c(0.01, 0.01), limits = c(-0.5, 30.5), n.breaks = 31) +
  labs(x = "Time difference to previous tracking ID of same class [s]",
       y = "Number of unique tracking IDs") +
  theme(legend.position = c(0.94, 0.8),
        plot.margin = margin(4, 2, 2, 2, "pt"))

```

```{r visit duration per class plot platform}
#plot of average visit duration (duration_s) per top1 class
insect_platform_cams <- platform_cams |>
  filter(str_detect(top1, "none", negate = TRUE))|>
  select(top1, duration_s) |>
   group_by(top1) |>
   summarise(
    mean_duration = mean(duration_s, na.rm = TRUE),
    sd_duration = sd(duration_s, na.rm = TRUE))


# Plot with error bars
ggplot(insect_platform_cams, aes(x = top1, y = mean_duration, fill = top1)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(
    ymin = mean_duration - sd_duration,
    ymax = mean_duration + sd_duration
  ), width = 0.2, color = "black") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, vjust = 0.3), legend.position = "none" ) +
  labs(title = "Platform visit duration per top1 class",
       x = "Insect Class",
       y = "Mean Duration (s)") +
  scale_fill_viridis_d()
```

```{r}
rm(timediff, insect_platform_cams, ct36, ct37, ct38, ct_all)
```

```{r save csv platform}
platform_cams_csv <- platform_cams |>
  #create a column combining site and transect
  mutate(Site_Tn= paste(location, transect, sep = "_")) |>
  #create a column combining camera ID and track_ID
  mutate(ID = paste(cam_ID, track_ID, sep = "_")) |>
  #remove "seppi-" prefix in ID
  mutate(ID = str_remove(ID, "seppi-")) |>
  select(Site_Tn, location, transect, date,  ID, cam_ID,start_time, top1, det_conf_mean, track_ID_imgs, top1_imgs, top1_prob_mean, top1_prob_weighted)

#save as csv
write_csv(platform_cams_csv, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/InsectDetect_platform_cams.csv")
rm(platform_cams_csv)
```

Sample_date	Site	camID	Transect_number
11/07/2024	WED	cam36	T3
14/07/2024	STP	cam36	T1
18/07/2024	JEP	cam36	T2
20/07/2024	KOT	cam36	T5
21/07/2024	DES	cam36	T1
24/07/2024	HLI	cam36	T1
29/07/2024	WUP	cam36	T3
30/07/2024	WDG	cam36	T5
13/08/2024	BUH	cam36	T1
11/07/2024	WED	cam37	T5
14/07/2024	STP	cam37	T5
18/07/2024	JEP	cam37	T5
20/07/2024	KOT	cam37	T1
21/07/2024	DES	cam37	T3
24/07/2024	HLI	cam37	T4
29/07/2024	WUP	cam37	T5
30/07/2024	WDG	cam37	T1
13/08/2024	BUH	cam37	T3
11/07/2024	WED	cam38	T2
14/07/2024	STP	cam38	T3
18/07/2024	JEP	cam38	T1
20/07/2024	KOT	cam38	T3
21/07/2024	DES	cam38	T5
24/07/2024	HLI	cam38	T5
29/07/2024	WUP	cam38	T1
30/07/2024	WDG	cam38	T3
13/08/2024	BUH	cam38	T5

```{r log_rec csv}
#on camptrap_logs_rec assign a site and transect to the recording logs, according to record start date and camera ID
camtrap_logs_rec_csv <- camtrap_logs_rec |>
  #site assignment
  mutate(Site = case_when(
    grepl("2024-07-11", record_start_date) ~ "WED",  # Match the 11th July to Wedringen
    grepl("2024-07-14", record_start_date) ~ "STP",  # Match the 14th July to Steutz
    grepl("2024-07-18", record_start_date) ~ "JEP",  # Match the 18th July to Jerichow
    grepl("2024-07-20", record_start_date) ~ "KOT",  # Match the 20th July to Köthen
    grepl("2024-07-21", record_start_date) ~ "DES",  # Match the 21nd July to Dessau
    grepl("2024-07-24", record_start_date) ~ "HLI",  # Match the 24th July to Pölsfeld
    grepl("2024-07-29", record_start_date) ~ "WUP",  # Match the 29th July to Wulfen
    grepl("2024-07-30", record_start_date) ~ "WDG",  # Match the 30th July to Wedringen
    grepl("2024-08-13", record_start_date) ~ "BUH",  # Match the 13th August to Bühne
    TRUE ~ "UNKNOWN"  # Default value
  )) |>
  #transect assignment
  mutate(Transect = case_when(
    grepl("seppi-cam36", cam_ID) & grepl("WED", Site) ~ "T3",  #cam36, WED, T3
    grepl("seppi-cam36", cam_ID) & grepl("STP", Site) ~ "T1",  #cam36, STP, T4
    grepl("seppi-cam36", cam_ID) & grepl("JEP", Site) ~ "T2",  #cam36, JEP, T2
    grepl("seppi-cam36", cam_ID) & grepl("KOT", Site) ~ "T5",  #cam36, KOT, T5
    grepl("seppi-cam36", cam_ID) & grepl("DES", Site) ~ "T1",  #cam36, DES, T1
    grepl("seppi-cam36", cam_ID) & grepl("HLI", Site) ~ "T1",  #cam36, HLI, T1
    grepl("seppi-cam36", cam_ID) & grepl("WUP", Site) ~ "T3",  #cam36, WUP, T3
    grepl("seppi-cam36", cam_ID) & grepl("WDG", Site) ~ "T5",  #cam36, WDG, T5
    grepl("seppi-cam36", cam_ID) & grepl("BUH", Site) ~ "T1",  #cam36, BUH, T1
    
    grepl("seppi-cam37", cam_ID) & grepl("WED", Site) ~ "T5",  #cam37, WED, T5
    grepl("seppi-cam37", cam_ID) & grepl("STP", Site) ~ "T5",  #cam37, STP, T5
    grepl("seppi-cam37", cam_ID) & grepl("JEP", Site) ~ "T5",  #cam37, JEP, T5
    grepl("seppi-cam37", cam_ID) & grepl("KOT", Site) ~ "T1",  #cam37, KOT, T1
    grepl("seppi-cam37", cam_ID) & grepl("DES", Site) ~ "T3",  #cam37, DES, T3
    grepl("seppi-cam37", cam_ID) & grepl("HLI", Site) ~ "T4",  #cam37, HLI, T4
    grepl("seppi-cam37", cam_ID) & grepl("WUP", Site) ~ "T5",  #cam37, WUP, T5
    grepl("seppi-cam37", cam_ID) & grepl("WDG", Site) ~ "T1",  #cam37, WDG, T1
    grepl("seppi-cam37", cam_ID) & grepl("BUH", Site) ~ "T3",  #cam37, BUH, T1

    grepl("seppi-cam38", cam_ID) & grepl("WED", Site) ~ "T2",  #cam38, WED, T2
    grepl("seppi-cam38", cam_ID) & grepl("STP", Site) ~ "T3",  #cam38, STP, T3
    grepl("seppi-cam38", cam_ID) & grepl("JEP", Site) ~ "T1",  #cam38, JEP, T1
    grepl("seppi-cam38", cam_ID) & grepl("KOT", Site) ~ "T3",  #cam38, KOT, T3
    grepl("seppi-cam38", cam_ID) & grepl("DES", Site) ~ "T5",  #cam38, DES, T5
    grepl("seppi-cam38", cam_ID) & grepl("HLI", Site) ~ "T5",  #cam38, HLI, T5
    grepl("seppi-cam38", cam_ID) & grepl("WUP", Site) ~ "T1",  #cam38, WUP, T1
    grepl("seppi-cam38", cam_ID) & grepl("WDG", Site) ~ "T3",  #cam38, WDG, T3
    grepl("seppi-cam38", cam_ID) & grepl("BUH", Site) ~ "T5",  #cam38, BUH, T5
    TRUE ~ "UNKNOWN"  # Default value
  )) |> 
  # Group by Site, Transect, and cam_ID
  group_by(Site, Transect, cam_ID) |> 
  summarise(
    # Sum the rec_time_min for each combination of Site, Transect, and cam_ID
    rec_time_min = sum(rec_time_min), 
    
    # Keep the first rec_start_time for each combination of Site, Transect, and cam_ID
    first_record_start_time = first(record_start_time),
    
    # Drop the grouping after summarization
    .groups = "drop"
  )
  
#save as csv
write_csv(camtrap_logs_rec_csv, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/platform_recording_logs.csv")
```


# Flower cameras


```{r load flower data}
camtrap31 <- read_csv("C:/Users/Almas/YOLOv5/yolov5-master/runs/predict-cls/seppi-cam31/results/2024-10-28_16-15-25_metadata_processed/seppi-cam31_metadata_classified_top1_final.csv", col_names = TRUE)
camtrap32 <- read_csv("C:/Users/Almas/YOLOv5/yolov5-master/runs/predict-cls/seppi-cam32/results/2024-10-28_16-38-45_metadata_processed/seppi-cam32_metadata_classified_top1_final.csv", col_names = TRUE)
camtrap33 <- read_csv("C:/Users/Almas/YOLOv5/yolov5-master/runs/predict-cls/seppi-cam33/results/2024-12-10_20-12-07_metadata_processed/seppi-cam33_metadata_classified_top1_final.csv", col_names = TRUE)
camtrap34 <- read_csv("C:/Users/Almas/YOLOv5/yolov5-master/runs/predict-cls/seppi-cam34/results/2024-12-10_20-12-37_metadata_processed/seppi-cam34_metadata_classified_top1_final.csv", col_names = TRUE)
camtrap35 <- read_csv("C:/Users/Almas/YOLOv5/yolov5-master/runs/predict-cls/seppi-cam35/results/2024-12-10_20-12-54_metadata_processed/seppi-cam35_metadata_classified_top1_final.csv", col_names = TRUE)

# Combine data from all camera traps
flower_cams <- bind_rows(
  camtrap31,
  camtrap32,
  camtrap33,
  camtrap34,
  camtrap35)

rm(camtrap31, camtrap32, camtrap33, camtrap34, camtrap35) # Remove individual cameras dataframes
```


```{r site assignment flowercam}
flower_cams <- flower_cams %>%
mutate(location = case_when(
    grepl("2024-07-11", start_time) ~ "WED",  # Match the 11th July to Wedringen
    grepl("2024-07-14", start_time) ~ "STP",  # Match the 14th July to Steutz
    grepl("2024-07-18", start_time) ~ "JEP",  # Match the 18th July to Jerichow
    grepl("2024-07-20", start_time) ~ "KOT",  # Match the 20th July to Köthen
    grepl("2024-07-21", start_time) ~ "DES",  # Match the 21nd July to Dessau
    grepl("2024-07-24", start_time) ~ "HLI",  # Match the 24th July to Pölsfeld
    grepl("2024-07-29", start_time) ~ "WUP",  # Match the 29th July to Wulfen
    grepl("2024-07-30", start_time) ~ "WDG",  # Match the 30th July to Wedringen
    grepl("2024-08-13", start_time) ~ "BUH",  # Match the 13th August to Bühne
    TRUE ~ "UNKNOWN"  # Default value
  ))
```

```{r flower assignment}
# Assign flower species to each camera trap
flower_cams <- flower_cams |>
  mutate(flower_sp = case_when(
    grepl("seppi-cam31", cam_ID) & grepl("2024-07-11", date) ~ "kna_arv", #cam31, 11th July, kna_arv
    grepl("seppi-cam31", cam_ID) & grepl("2024-07-14", date) ~ "dau_car", #cam31, 14th July, dau_car
    grepl("seppi-cam31", cam_ID) & grepl("2024-07-18", date) ~ "hyp_per", #cam31, 18th July, hyp_per
    grepl("seppi-cam31", cam_ID) & grepl("2024-07-20", date) ~ "pas_sat", #cam31, 20th July, pas_sat
    grepl("seppi-cam31", cam_ID) & grepl("2024-07-21", date) ~ "dau_car", #cam31, 21st July, dau_car
    grepl("seppi-cam31", cam_ID) & grepl("2024-07-24", date) ~ "pas_sat", #cam31, 24th July, pas_sat  
    grepl("seppi-cam31", cam_ID) & grepl("2024-07-29", date) ~ "pas_sat", #cam31, 29th July, pas_sat
    grepl("seppi-cam31", cam_ID) & grepl("2024-07-30", date) ~ "gal_alb", #cam31, 30th July, gal_alb
    grepl("seppi-cam31", cam_ID) & grepl("2024-08-13", date) ~ "lot_cor", #cam31, 13th August, lot_cor
    
    grepl("seppi-cam32", cam_ID) & grepl("2024-07-11", date) ~ "kna_arv", #cam32, 11th July, kna_arv
    grepl("seppi-cam32", cam_ID) & grepl("2024-07-14", date) ~ "cre_bie", #cam32, 14th July, cre_bie
    grepl("seppi-cam32", cam_ID) & grepl("2024-07-18", date) ~ "peu_off", #cam32, 18th July, peu_off
    grepl("seppi-cam32", cam_ID) & grepl("2024-07-20", date) ~ "kna_arv", #cam32, 20th July, kna_arv
    grepl("seppi-cam32", cam_ID) & grepl("2024-07-21", date) ~ "cre_bie", #cam32, 21st July, cre_bie
    grepl("seppi-cam32", cam_ID) & grepl("2024-07-24", date) ~ "kna_arv", #cam32, 24th July, kna_arv
    grepl("seppi-cam32", cam_ID) & grepl("2024-07-29", date) ~ "cre_bie", #cam32, 29th July, cre_bie
    grepl("seppi-cam32", cam_ID) & grepl("2024-07-30", date) ~ "pas_sat", #cam32, 30th July, pas_sat
    grepl("seppi-cam32", cam_ID) & grepl("2024-08-13", date) ~ "gal_alb", #cam32, 13th August, gal_alb
    
    grepl("seppi-cam33", cam_ID) & grepl("2024-07-11", date) ~ "dau_car", #cam33, 11th July, dau_car
    grepl("seppi-cam33", cam_ID) & grepl("2024-07-14", date) ~ "tri_rep", #cam33, 14th July, tri_rep
    grepl("seppi-cam33", cam_ID) & grepl("2024-07-18", date) ~ "gal_bor", #cam33, 18th July, gal_bor
    grepl("seppi-cam33", cam_ID) & grepl("2024-07-20", date) ~ "cre_bie", #cam33, 20th July, cre_bie
    grepl("seppi-cam33", cam_ID) & grepl("2024-07-21", date) ~ "cre_bie", #cam33, 21st July, cre_bie
    grepl("seppi-cam33", cam_ID) & grepl("2024-07-24", date) ~ "hyp_per", #cam33, 24th July, hyp_per
    grepl("seppi-cam33", cam_ID) & grepl("2024-07-29", date) ~ "pas_sat", #cam33, 29th July, pas_sat
    grepl("seppi-cam33", cam_ID) & grepl("2024-07-30", date) ~ "cen_jac", #cam33, 30th July, cen_jac
    grepl("seppi-cam33", cam_ID) & grepl("2024-08-13", date) ~ "tri_pra", #cam33, 13th August, tri_pra
    
    grepl("seppi-cam34", cam_ID) & grepl("2024-07-11", date) ~ "pas_sat", #cam34, 11th July, pas_sat
    grepl("seppi-cam34", cam_ID) & grepl("2024-07-14", date) ~ "kna_arv", #cam34, 14th July, kna_arv
    grepl("seppi-cam34", cam_ID) & grepl("2024-07-18", date) ~ "peu_off", #cam34, 18th July, peu_off
    grepl("seppi-cam34", cam_ID) & grepl("2024-07-20", date) ~ "dau_car", #cam34, 20th July, dau_car
    grepl("seppi-cam34", cam_ID) & grepl("2024-07-21", date) ~ "tri_pra", #cam34, 21st July, tri_pra
    grepl("seppi-cam34", cam_ID) & grepl("2024-07-24", date) ~ "tri_pra", #cam34, 24th July, tri_pra
    grepl("seppi-cam34", cam_ID) & grepl("2024-07-29", date) ~ "dau_car", #cam34, 29th July, dau_car
    grepl("seppi-cam34", cam_ID) & grepl("2024-07-30", date) ~ "dia_car", #cam34, 30th July, dia_car
    grepl("seppi-cam34", cam_ID) & grepl("2024-08-13", date) ~ "dau_car", #cam34, 13th August, dau_car
    
    grepl("seppi-cam35", cam_ID) & grepl("2024-07-11", date) ~ "pas_sat", #cam35, 11th July, pas_sat
    grepl("seppi-cam35", cam_ID) & grepl("2024-07-14", date) ~ "hyp_per", #cam35, 14th July, hyp_per
    grepl("seppi-cam35", cam_ID) & grepl("2024-07-18", date) ~ "pas_sat", #cam35, 18th July, pas_sat
    grepl("seppi-cam35", cam_ID) & grepl("2024-07-20", date) ~ "cre_bie", #cam35, 20th July, cre_bie
    grepl("seppi-cam35", cam_ID) & grepl("2024-07-21", date) ~ "pas_sat", #cam35, 21st July, pas_sat
    grepl("seppi-cam35", cam_ID) & grepl("2024-07-24", date) ~ "hyp_per", #cam35, 24th July, hyp_per
    grepl("seppi-cam35", cam_ID) & grepl("2024-07-29", date) ~ "pas_sat", #cam35, 29th July, pas_sat
    grepl("seppi-cam35", cam_ID) & grepl("2024-07-30", date) ~ "dau_car", #cam35, 30th July, dau_car
    grepl("seppi-cam35", cam_ID) & grepl("2024-08-13", date) ~ "tri_pra" #cam35, 13th August, tri_pra
  ))

```

```{r load flower recording logs flowercam}

camtrap31_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam31/record_log.csv", col_names=T) |> mutate(
    # Extract date and time components from rec_start and rec_end
    record_start_date = as_date(rec_start), # Convert to Date
    record_start_time = format(rec_start, "%H:%M"), # Extract Time as HH:MM
    record_end_time = format(rec_end, "%H:%M")) # Extract Time as HH:MM

camtrap32_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam32/record_log.csv", col_names=T) |> mutate(
    # Extract date and time components from rec_start and rec_end
    record_start_date = as_date(rec_start), # Convert to Date
    record_start_time = format(rec_start, "%H:%M"), # Extract Time as HH:MM
    record_end_time = format(rec_end, "%H:%M") # Extract Time as HH:MM
  )

camtrap33_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam33/record_log.csv", col_names=T) |> mutate(
    # Extract date and time components from rec_start and rec_end
    record_start_date = as_date(rec_start), # Convert to Date
    record_start_time = format(rec_start, "%H:%M"), # Extract Time as HH:MM
    record_end_time = format(rec_end, "%H:%M") # Extract Time as HH:MM
  )

camtrap34_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam34/record_log.csv", col_names=T) |> mutate(
    # Extract date and time components from rec_start and rec_end
    record_start_date = as_date(rec_start), # Convert to Date
    record_start_time = format(rec_start, "%H:%M"), # Extract Time as HH:MM
    record_end_time = format(rec_end, "%H:%M") # Extract Time as HH:MM
  )

camtrap35_logs_rec <-  read_csv("C:/Users/Almas/Desktop/Alma/seppi-cam35/record_log.csv", col_names=T) |> mutate(
    # Extract date and time components from rec_start and rec_end
    record_start_date = as_date(rec_start), # Convert to Date
    record_start_time = format(rec_start, "%H:%M"), # Extract Time as HH:MM
    record_end_time = format(rec_end, "%H:%M") # Extract Time as HH:MM
  )


flower_logs_rec <- bind_rows(camtrap31_logs_rec, camtrap32_logs_rec, camtrap33_logs_rec, camtrap34_logs_rec, camtrap35_logs_rec)
rm(camtrap31_logs_rec, camtrap32_logs_rec, camtrap33_logs_rec, camtrap34_logs_rec, camtrap35_logs_rec)
```

## 6 Flower camera data summary

```{r data summary flowercam}
# Total recording time of all camera traps
sum(flower_logs_rec$rec_time_min) / 60
# ~251.6585h

# Total recording time per camera trap
flower_logs_rec |> 
  group_by(cam_ID) |> 
  summarise(record_time_total_h = sum(rec_time_min) / 60)
#seppi-cam31	54.00950			
#seppi-cam32	54.00950			
#seppi-cam33	53.87150			
#seppi-cam34	53.89483			
#seppi-cam35	35.87317	

# Total number of captured images
sum(flower_cams$track_ID_imgs)
# 192 290 images


# Total number of all unique tracking IDs
nrow(flower_cams)
# 5553  tracking IDs total

# Total number of all unique tracking IDs of insect classes
flower_cams |> 
  filter(str_detect(top1, "none", negate = TRUE)) |> # how many insect pictures are left when excluding the none_something classes
  nrow()
# 445 tracking IDs of insect classes

# Total number of all unique tracking IDs of non-insect classes
flower_cams |> 
  filter(str_detect(top1, "none")) |> # how many none_something pictures are there
  nrow()
# 5105 tracking IDs of non-insect classes (background, bird, dirt, shadow)

  
# Total number of captured images, for which the top1 class with the highest weighted probability was hoverfly
flower_cams |> 
  filter(str_detect(top1, "hfly")) |>
  summarise(track_ID_imgs_hfly = sum(track_ID_imgs))
# 784 images of hoverfly classes

# Total number of all unique tracking IDs that were classified as hoverfly
flower_cams |> 
  filter(str_detect(top1, "hfly")) |>
  nrow()
# 35 tracking IDs of hoverfly classes

# Total number of all unique tracking IDs per hoverfly class
flower_cams |> 
  filter(str_detect(top1, "hfly")) |>
  count(top1)
#hfly_episyr	3			
#hfly_eristal	12			
#hfly_eupeo	16			
#hfly_sphaero	4	

table(flower_cams$top1)
```

## 7 Flower camera plotting
```{r recording duration per session flowercam}
#visualize the recording duration per session
ggplot(flower_logs_rec, aes(x = cam_ID, y = rec_time_min/60, #divide by 60 to get hours
                             fill=rec_time_min/60)) + #fill the bars with the total hours recorded
  geom_bar(stat= "identity") +
  geom_col(colour = "grey")+ #add grey outline to the stacked sections
  labs(title = "Recording duration per session", # Session not day bc some recordings are split within one day
       x = "Camera",
       y = "Total hours recorded") +
  labs(fill = "Hours recorded") +
  theme_minimal()+
  scale_fill_viridis(direction = -1) #reverse the viridis color scale
```

```{r unique IDs per class flowercam}
# Generate color palette to plot the same color for each site respectively,
# even if the number of months the camera traps were active differs
viridis_9 <- viridis_pal()(9) #number of sites

(ct31 <- flower_cams |>
  filter(str_detect(cam_ID, "cam31")) |>
  barplot_total(class_exclude = "none", title = "Camtrap 31") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 15), n.breaks = 16))

(ct32 <- flower_cams |>
  filter(str_detect(cam_ID, "cam32")) |>
  barplot_total(class_exclude = "none", title = "Camtrap 32") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 40), n.breaks = 16))

(ct33 <- flower_cams |>
  filter(str_detect(cam_ID, "cam33")) |>
  barplot_total(class_exclude = "none", title = "Camtrap 33") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 50), n.breaks = 16))

(ct34 <- flower_cams |>
    filter(str_detect(cam_ID, "cam34")) |>
    barplot_total(class_exclude = "none", title = "Camtrap 34") +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 20), n.breaks = 16))

(ct35 <- flower_cams |>
    filter(str_detect(cam_ID, "cam35")) |>
    barplot_total(class_exclude = "none", title = "Camtrap 35") +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 100), n.breaks = 16))

(ct_all_flower <- flower_cams |>
  barplot_total(class_exclude = "none", title = "All camera traps") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 200), n.breaks = 16))
  
# Combine plots
(ct31 + labs(x = "")) +
(ct32 + labs(x = "", y = "")) +
(ct33 + labs(x = "", y = "")) +
(ct34 + labs(x = "", y = "")) +
(ct35 + labs(x = "", y = "")) +
(ct_all_flower + labs(y = "")) +
  plot_annotation(tag_levels = "A") & 
  theme(plot.tag = element_text(size = 11, face = "bold"))

# Plot all classes, including "none_bg", "none_bird", "none_dirt", "none_shadow"
flower_cams |> 
  barplot_total(class_exclude = "0", title = "Camtraps merged") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 5000), n.breaks = 14)

flower_cams |> 
  filter(str_detect(top1, "none",  negate = TRUE)) |> # Exclude none_ class
  ggplot(aes(x = location, fill = top1)) +
  geom_bar() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3)) +
  labs(title = "Number of unique tracking IDs per site",
       x = "Site",
       y = "Number of unique tracking IDs") +
  scale_fill_viridis_d()+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 300))

```

```{r visit duration per class flowercam}
#plot of average visit duration (duration_s) per top1 class
insect_flower_cams <- flower_cams |>
  filter(str_detect(top1, "none", negate = TRUE))|>
  select(top1, duration_s) |>
   group_by(top1) |>
   summarise(
    mean_duration = mean(duration_s, na.rm = TRUE),
    sd_duration = sd(duration_s, na.rm = TRUE))


# Plot with error bars
ggplot(insect_flower_cams, aes(x = top1, y = mean_duration, fill = top1)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(
    ymin = mean_duration - sd_duration,
    ymax = mean_duration + sd_duration
  ), width = 0.2, color = "black") +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1, vjust = 0.3), legend.position = "none" ) +
  labs(title = "Platform visit duration per top1 class",
       x = "Insect Class",
       y = "Mean Duration (s)") +
  scale_fill_viridis_d()

```

### saving dataframe for further analysis
```{r save dataframes}
current_date <- format(Sys.Date(), "%Y%m%d")

# save flower_cams 
#remove all none_ categories
flower_cams_save <- flower_cams |>
  filter(str_detect(top1, "none", negate = TRUE)) |>
  mutate(site_type = case_when( #adding site_type column
                location %in% c("WED", "KOT", "WDG", "BUH") ~ "young_restored", #young restored sites
                location %in% c("STP", "JEP", "DES", "HLI", "WUP") ~ "reference", #reference sites
                TRUE ~ NA_character_))  |> # Default case, if needed
  rename("site" = "location") #rename location to site
#save flower_cams as RData file
save(flower_cams_save, file = paste0("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/", current_date,"_flower_cams.RData"))

# save platform_cams
#remove all none_ categories
platform_cams_save <- platform_cams |>
  filter(str_detect(top1, "none", negate = TRUE)) |>
  mutate(site_type = case_when( #adding site_type column
                location %in% c("WED", "KOT", "WDG", "BUH") ~ "young_restored", #young restored sites
                location %in% c("STP", "JEP", "DES", "HLI", "WUP") ~ "reference", #reference sites
                TRUE ~ NA_character_))  |> # Default case, if needed
  rename("site" = "location") #rename location to site
#save platform_cams as RData file
save(platform_cams_save, file = paste0("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/", current_date,"_platform_cams.RData"))
```