---
title: "Landscape"
author: "Alma Steireif"
date: "`r Sys.Date()`"
output: html_document
---


Author: Alma Steireif
Date: `r Sys.Date()`

### Description

The goal of this script is to create a .csv file that contains the landscape data extracted from CORINE 2018 and ESA WorldCover V2 2021 around the 9 study sites. 

Data was accessed from the Copernicus Land Monitoring Service (CLMS) and the European Space Agency (ESA).
- CORINE 2018 : https://land.copernicus.eu/en/products/corine-land-cover/clc2018#download
- WorldCover V2 2021: https://esa-worldcover.org/en

The data was then processed in QGIS to extract the surface in square meters covered by each land cover class in 1km buffers around the study sites. The data was then exported as .csv files. 

For the purpose of this research, we kept only the CORINE land cover classes, adn combined some of the classes to reduce the number of classes in the data. We also calculated the relative cover ratio for each land cover class in each buffer. 

```{r setup, include=FALSE}
library(tidyverse)
```

### load data

```{r}
# load CORINE from "C:\Users\Almas\Desktop\UNI_LEIPSI\Thesis\Thesis_Rproject\Landscape\corine_area_by_buffer.csv"
corine <- read.csv("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/Landscape/corine_area_by_buffer.csv")

#load WorldCover from "C:\Users\Almas\Desktop\UNI_LEIPSI\Thesis\Thesis_Rproject\Landscape\ESAWC_landcover_area_by_buffer.csv"
worldcover <- read.csv("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/Landscape/ESAWC_landcover_area_by_buffer.csv")
``` 


### Data cleaning

#### CORINE

The landcover classes are coded as numbers in the data. For now we are keeping the codes as they are, and adding a prefix "CORINE" to the landcover class names.

| Code | Landcover class |
|:-----|:----------------|
| 01    | Continuous urban fabric |
| 02    | Discontinuous urban fabric |
| 03    | Industrial or commercial units |
| 11   | Sports and leisure facilities |
| 12   | Non-irrigated arable land |
| 18   | Pastures |
| 23   | Broad-leaved forest |
| 24   | Coniferous forest |
| 26   | Natural grasslands |
| 29   | Transitional woodland-shrub |
| 35   | Inland marshes |

The data is in a long format, with each row representing a land cover class in a buffer around a study site. We will pivot the data to a wide format, with each row representing a study site and each column representing a land cover class. 

```{r CORINE}
#rename "Buffer" to "Site"
colnames(corine)[1] <- "Site"

#rename the rows in LandCover equal to 1,2 and3 with 01,02 and 03
corine$LandCover <- ifelse(corine$LandCover < 10, paste0("0", corine$LandCover), corine$LandCover)

#rename the rows in LandCover with the prefix CORINE
#corine$LandCover <- paste("CORINE", corine$LandCover, sep = "_")

# pivot the data to wide format
corine_wide <- corine %>% 
  #rename the rows in LandCover with the prefix CORINE
  mutate(LandCover = paste("CORINE", LandCover, sep = "_")) %>%
  pivot_wider(names_from = LandCover, values_from = Area_m2)%>%
  #replace NA with 0
  replace(is.na(.), 0)
```

The CORINE dataset has many classes of land cover, some of which are very similar and this level of precision is not needed in the scope of this research. We will combine some of the classes to reduce the number of classes in the data. Here are the new 6 classes

1. Agricultural land - agri
    - Non-irrigated arable land
    
2. Grassland - grass
    - Pastures
    - Natural grasslands
3. Semi-natural habitat - snh
    - transitional woodland-shrub
4. Forest - forest
    - Broad-leaved forest
    - Coniferous forest
5. Urban - urban
    - Continuous urban fabric
    - Discontinuous urban fabric
    - Industrial or commercial units
    - Sports and leisure facilities
6. Water - water
    - Inland marshes

```{r combine classes}
#combine the classes
corine_wide_reduced <- corine_wide %>%
  mutate(agri = CORINE_12,
         grass = CORINE_18 + CORINE_26,
         snh = CORINE_29,
         forest = CORINE_23 + CORINE_24,
         urban = CORINE_01 + CORINE_02 + CORINE_03 + CORINE_11,
         water = CORINE_35) %>%
  select(Site, agri, grass, snh, forest, urban, water)
```

##### CORINE - Relative cover ratio

For now, the data is in square meters. We will calculate the relative cover ratio for each land cover class in each buffer. The relative cover ratio is the proportion of the buffer that is covered by each land cover class. 

$$Relative\ cover\ ratio = \frac{Area\ of\ landcover\ class}{Total\ area\ of\ buffer}$$
Since the buffers are all 1km in radius, the total area of the buffer is $pi*r^2= pi*(1000m)^2 = 3141592.65359 m^2$


```{r relative_cover_ratio}
#calculate the relative cover ratio
corine_relative <- corine_wide_reduced %>%
  mutate(agri = agri / 3141592.65359,
         grass = grass / 3141592.65359,
         snh = snh / 3141592.65359,
         forest = forest / 3141592.65359,
         urban = urban / 3141592.65359,
         water = water / 3141592.65359)%>%
  #times 100 to get the percentage
  mutate(across(agri:water, ~ .x * 100))%>%
    #round to 4 decimal places
  mutate(across(agri:water, ~ round(.x, 3)))
```


Since we have a maximum of 45 sampling points over the whole experience, we can not have too many explanatory variables. We will also create a new catgorical column that contains the class which occupied the majority of the buffer. 

```{r}
# create a new column with the class which occupied the majority of the buffer
corine_relative <- corine_relative %>%
  mutate(majority_class = case_when(
    agri > grass & agri > snh & agri > forest & agri > urban & agri > water ~ "agri",
    grass > agri & grass > snh & grass > forest & grass > urban & grass > water ~ "grass",
    snh > agri & snh > grass & snh > forest & snh > urban & snh > water ~ "snh",
    forest > agri & forest > grass & forest > snh & forest > urban & forest > water ~ "forest",
    urban > agri & urban > grass & urban > snh & urban > forest & urban > water ~ "urban",
    water > agri & water > grass & water > snh & water > forest & water > urban ~ "water"
  ))
```



### Save the data

```{r}
write.csv(corine_relative, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/corine_data.csv", row.names = FALSE)
```

#### ESA WorldCover

The landcover classes are also coded as numbers in the data. We will add a prefix "ESAWC" to the landcover class names and transform the data to wide format.

| Code | Landcover class |
|:-----|:----------------|
| 10   | Tree Cover |
| 30   | Grassland |
| 40   | Cropland |
| 50   | Built-up |
| 60   | Bare / Sparse vegetation |
| 80   | Permanent Water Bodies |
| 90   | Wetlands |


```{r ESAWC}
#rename "Buffer" to "Site"
colnames(worldcover)[1] <- "Site"

#rename the rows in LandCover with the prefix ESAWC
worldcover$LandCover <- paste("ESAWC", worldcover$LandCover, sep = "_")

# pivot the data to wide format
worldcover_wide <- worldcover %>% 
  pivot_wider(names_from = LandCover, values_from = Area_m2)%>%
  #replace NA with 0
  replace(is.na(.), 0)

#site names - add a prefix indicating the site_type
#  for "DES", "HLI", "JEP", "STP", "WUP"  add "ref" - reference sites
# for  "BUH" ,   "KOT"  ,  "WDG",    "WED" add "res" - restored sites

  #add site_type prefix
worldcover_wide$Site <- ifelse(worldcover_wide$Site %in% c("DES", "HLI", "JEP", "STP", "WUP"), paste("ref", worldcover_wide$Site, sep = "_"), paste("res", worldcover_wide$Site, sep = "_"))
corine_wide$Site <- ifelse(corine_wide$Site %in% c("DES", "HLI", "JEP", "STP", "WUP"), paste("ref", corine_wide$Site, sep = "_"), paste("res", corine_wide$Site, sep = "_"))

```


### Vizualise the data

```{r}
#vector continaing the CORINE legend for the pie chart with the landcover classes names
CORINE_legend <- c("Continuous urban fabric", "Discontinuous urban fabric", "Industrial or commercial units", "Sports and leisure facilities", "Non-irrigated arable land", "Pastures", "Broad-leaved forest", "Coniferous forest", "Natural grasslands", "Transitional woodland-shrub", "Inland marshes")

#plot the data in a pie chart for CORINE only with custom legend
corine_wide %>%
  gather(key = "LandCover", value = "Area_m2", -Site) %>%
  ggplot(aes(x = "", y = Area_m2, fill = LandCover)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  facet_wrap(~Site) +
  theme_void() +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("CORINE_01" = "#FA6900FF", "CORINE_02" = "#F38630FF", "CORINE_03" = "#BE3428FF", "CORINE_11" = "#8785B2FF", "CORINE_12" = "#D6CFB7FF", "CORINE_18" = "#a3b18a", "CORINE_23" = "#121510FF", "CORINE_24" = "#006400", "CORINE_26" = "#6D8325FF", "CORINE_29" = "#647D4BFF", "CORINE_35" = "#7FC7AFFF"), labels = CORINE_legend) +
  labs(title = "2018 CORINE Landcover in 1km buffers around study sites")

#plot the data in a pie chart for CORINE only with custom legend
plot1 <- corine_wide_reduced %>%
  gather(key = "LandCover", value = "Area_m2", -Site) %>%
  ggplot(aes(x = "", y = Area_m2, fill = LandCover)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  facet_wrap(~Site) +
  theme_void() +
  theme(legend.position = "right") +
  scale_fill_manual(values = c("agri" = "#D6CFB7FF", "grass" = "#6D8325FF", "snh" = "#647D4BFF", "forest" = "#121510FF", "urban" = "#FA6900FF", "water" = "#7FC7AFFF")) +
  labs(title = "2018 CORINE Landcover in 1km \t
       buffers around study sites")

#vector containaing the ESAWC legend for the pie chart with the landcover classes names
ESAWC_legend <- c("Tree Cover", "Grassland", "Cropland", "Built-up", "Bare / Sparse vegetation", "Permanent Water Bodies", "Wetlands")

#plot the data in a pie chart for ESAWC only with legend
plot2 <- worldcover_wide %>%
  gather(key = "LandCover", value = "Area_m2", -Site) %>%
  ggplot(aes(x = "", y = Area_m2, fill = LandCover)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  facet_wrap(~Site) +
  theme_void() +
  theme(legend.position = "right") +
  scale_fill_manual(values = c("ESAWC_10" = "#121510FF", "ESAWC_30" = "#6D8325FF", "ESAWC_40" = "#D6CFB7FF", "ESAWC_50" = "#BD5630FF", "ESAWC_60" = "#647D4BFF", "ESAWC_80" = "#3FB8AFFF", "ESAWC_90" = "#7FC7AFFF"), labels = ESAWC_legend) +
  labs(title = "2021 ESAWC Landcover in 1km \t
       buffers around study sites")

#combine the plots and adjust the margins
final_plot <- plot1 | plot2 + plot_layout(guides = "collect") & 
  theme(plot.margin = margin(5, 5, 5, 5)) # Adjust margins if needed

print(final_plot)
```
