---
title: "plant_diversity"
author: "Alma Steireif"
date: "`r Sys.Date()`"
output: html_document
---

# 1. Set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(googlesheets4) #package to read google sheets
library(tidyverse) 
library(viridis) #color palette

library(iNEXT) #package to run rarefaction curves
library(fossil) 

library(vegan)#package to run metaNMDS

library(TNRS) # Resolve plant taxonomic names

plant_data_load <-read_sheet("https://docs.google.com/spreadsheets/d/1YDOoUobRU6A36um0-iut6kY77X0jEMYXtFWR3-Zankc/edit?usp=sharing", sheet=3) #sheet 3 has the plant data
```

```{r custom colors}

#define custom colors for plotting
custom_colors <- c(
  #reference sites
  "DES" = "#FFCC00",  # Yellow
  "HLI" = "#FF9900",  # Dark Orange
  "JEP" = "#FF6600",  # Deep Orange
  "STP" = "#B53737FF",  # Dark Orange
  "WUP" = "#5E0000FF",  # Red-Orange
  # restored sites
  "BUH" = "#006400",  # Dark Green
  "KOT" = "#5aae61",  # Medium Green
  "WDG" = "#67a9cf",  # Light Blue
  "WED" = "#4682B4"   # Steel Blue
)

#pastel palette for confidence intervals
pastel_colors <- c(
  # reference sites 
  "DES" = "#FDF3B9",  # Pale Yellow
  "HLI" = "#F6D6AD",  # Light Peach
  "JEP" = "#F4C095",  # Pastel Orange
  "STP" = "#F7CAC9",  # Soft Coral
  "WUP" = "#EDA9ABFF",  # 
  #restored sites 
  "BUH" = "#A8D5BA",  # Pastel Green
  "KOT" = "#B8E0D2",  # Pastel Teal
  "WDG" = "#A7C7E7",  # Light Blue
  "WED" = "#89ABE3"   # Soft Periwinkle
)

# bicolor palette
bicolor_palette <- c(
  "DES" = "darkorange",  
  "HLI" = "darkorange",  
  "JEP" = "darkorange",  
  "STP" = "darkorange",
  "WUP" = "darkorange",
  "BUH" = "steelblue",  
  "KOT" = "steelblue",  
  "WDG" = "steelblue",  
  "WED" = "steelblue" 
)

bicolor_bg <- c(
  "DES" = "lightyellow",  
  "HLI" = "lightyellow",  
  "JEP" = "lightyellow",  
  "STP" = "lightyellow",
  "WUP" = "lightyellow",
  "BUH" = "lightblue1",  
  "KOT" = "lightblue1",  
  "WDG" = "lightblue1",  
  "WED" = "lightblue1" 
)

```


# 2. Data preparation
## 2.1. Selecting columns & cleaning 
```{r selecting columns & cleaning}
plants <- plant_data_load %>%
  select(1:7) %>%
  
  # Rename columns for consistency
  rename(Site_type = Age,  # Change "Age" to "Site_type" for standardization with other scripts
    Plant_sp = `plant species`, 
    Cover_percentage = `cover (%)`, 
    Transect= transect,
    Plot=plot) %>%
  
  select(-notes)%>% #remove notes column
  filter(Plant_sp != "no flowers")%>% #remove "no flowers" rows in Plant_sp col

  #rm NA in Site
  filter(!is.na(Site))

head(plants)
```
## 2.2. Standardize plant names
```{r standardize plant names TNRS}
colnames(plants)
#next step requires internet connection
out_rom<- TNRS(taxonomic_names = plants$Plant_sp) # indicating where the scientific names to verify are
head(out_rom[,2:7])

to_correct <-out_rom %>% filter(Name_score < 1)
print(to_correct);length(to_correct$Name_score) #print the names that need to be corrected, in this case we have 0
#plants$plant<-gsub("previous name","correct name", plants$Plant_species) #replace the previous name with the correct name in a new column

#removing variables that are not needed for the next steps
rm(out_rom, to_correct)
```


## 2.3. Fixing Flower cover estimates (Braun-Blanquet cover abundance scale)
- r = rare (1-3 individuals), cover up to 0.5% of the plot [0.1-0.5%]   
- +  = 3-5 individuals, cover up to 1% of the plot [0.5-1%]   
- 1 = many individuals (6 to 50 individuals), cover up to 5% of the plot [1-5%]   
- 2 = a lot (>50 individuals), cover up between 5 and 25% of the plot [5-25%]   
- 3 = very abundant, cover up to 50% of the plot [25-50%]   
- 4 = dominant, cover up to 75% of the plot [50-75%]   
- 5 = very dominant, cover up to 100% of the plot [75-100%]  

```{r braun-blanquet}
#consistency in structure for cover_percentage
plants <- plants %>%
  #convert cover_percentage to character
  mutate(Cover_percentage = as.character(Cover_percentage)) %>%
  #transform "NA" to NA
  mutate(Cover_percentage = ifelse(Cover_percentage == "NA", NA, Cover_percentage))
    
#create a new column for BB scale and fill it
plants3 <- plants %>%
  #New column for BB scale where only "r", "+", 0 and NA are moved for now
  mutate(BB=case_when(
    Cover_percentage == "r" ~ "r",
    Cover_percentage == "+" ~ "+",
    Cover_percentage == "0" ~ "0",
    is.na(Cover_percentage) ~ NA_character_,
    #rest is empty for now
    TRUE ~ "")) %>%
  #remove "r", "+", 0 from Cover_percentage
  mutate(Cover_percentage = case_when(
    Cover_percentage == "r" ~ NA_character_,
    Cover_percentage == "+" ~ NA_character_,
    Cover_percentage == "0" ~ NA_character_,
    TRUE ~ Cover_percentage)) %>%
  #convert Cover_percentage to numeric
  mutate(Cover_percentage = as.numeric(Cover_percentage)) %>%
  #fill in the empty BB values based on BB scale
  mutate(BB = case_when(
    BB=="" & Cover_percentage <= 0.5 ~ "r",
    BB=="" & Cover_percentage < 1 ~ "+",
    BB=="" & Cover_percentage < 5 ~ "1",
    BB=="" & Cover_percentage < 25 ~ "2",
    BB=="" & Cover_percentage < 50 ~ "3",
    BB=="" & Cover_percentage < 75 ~ "4",
    TRUE ~ BB))

# fill in the NA values in Cover_percentage based on BB scale
plants3 <- plants3 %>%
  mutate(Cover_percentage = case_when(
    is.na(Cover_percentage) & BB == "r" ~ 0.3,
    is.na(Cover_percentage) & BB == "+" ~ 0.75,
    TRUE ~ Cover_percentage))%>%
  #BB as factor
  mutate(BB = factor(BB, levels = c("r", "+", "1", "2", "3", "4", "5")))
  
rm(plant_data_load, plants)
head(plants3)
```


```{r braun-blanquet plot}
#barplot of coverage per Site
ggplot(plants3, aes(x = Site, y = Cover_percentage, fill= Plant_sp)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Plant Species Coverage",
    x = "Site",
    y = "Cover Percentage (%)",
    fill = "Braun-Blanquet Scale"
  )+
  #remove legend
  theme(legend.position = "none")+
  # adding black outline to bars
  geom_bar(stat = "identity", color = "black")+
  #viridis color palette
  scale_fill_viridis(discrete = TRUE)
```
# 3. Analysis

## A. Richness calculation
```{r}
richness <- plants3 %>%
  group_by(Site, Site_type) %>%
  summarise(Species_Richness = n_distinct(Plant_sp))

# point plot richness
ggplot(richness, aes(x = Site, y = Species_Richness, color = Site_type)) +
  geom_point(size = 3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Plant Species Richness",
    x = "Site",
    y = "Number of Flower Species",
    color = "Site Type")
```


## B.1. Hill number calculation (not pooled) - presence/absence at plot level 

*Species-by-sampling-unit matrix (like a checklist for each subplot)*

- Convert the flower cover percentage cover data to presence-absence data per plot (1 = present, 0 = absent). There were 5 transects per site, and each transect had three plots. 
- Replace NA values in the Cover_percentage column with 1 (presence), because since the species was written down, it means that it was present in the plot, but the cover % was forgotten/not written down properly. 
- Remove the "no flowers" rows in the Plant_sp column, as they do not contribute to species richness.
- Convert it to matrix format for iNEXT analysis
- Add top row with the total number of species per site in order to use  datatype = `'incidence_freq'`` 

- $AsyEst lists the observed diversity, asymptotic estimates, estimated bootstrap s.e. and 95% (default) confidence intervals for Hill numbers of order q = 0, 1, and 2. Theestimated asymptotes are calculated via the functions ChaoRichness() for q=0, ChaoShannon() for q=1 and ChaoSimpson() for q=2; see Chao et al. (2014) for the formulas of these asymptotic estimators. [https://drive.google.com/file/d/1Z9T6xMD7IOSo1SqyF2BQou7Fez5evlhj/view] 
 
```{r preparing for iNEXT}
# Presence Absence transformation of Cover_percentage
plants1 <- plants3 %>%
  #transform plot column into factor
  mutate(Plot = factor(Plot))%>%
  #new presence/absence column
  mutate(Presence_Absence = ifelse(Cover_percentage > 0, 1, 0))%>%
  #Replace Na rows in Presence_Absence with 1 (presence)
  mutate(Presence_Absence = ifelse(is.na(Presence_Absence), 1, Presence_Absence))

plants1 <- as.data.frame(plants1)

# Create a matrix for iNEXT
matrix_plant <- create.matrix(x = plants1,
                                tax.name = 'Plant_sp',
                                locality = 'Site',
                                abund = TRUE,
                                abund.col = 'Presence_Absence')
#Check your work
identical(sum(matrix_plant),sum(plants1$Presence_Absence))
#both have 646 observations

# Convert to data.frame for use with iNEXT
matrix_plant <- as.data.frame(matrix_plant)

# Assign column names (site names)
colnames(matrix_plant) <- c("BUH", "DES", "HLI", "JEP", "KOT", "STP", "WDG", "WED", "WUP")

# Assign row names (species names) = unique and ordered plant species from Plant_sp column
rownames(matrix_plant) <- plants1$Plant_sp %>% unique() %>% sort()

# Define the number of sampling units per site 
total_units <- rep(15, ncol(matrix_plant))  # 15 plots per site

# Add as a new first row in the matrix
matrix_plant <- rbind(Total_units = total_units, matrix_plant)

# Print the updated matrix
head(matrix_plant)
```

```{r Hill number calculation}
#run iNEXT to get the diversity estimates per SITE_TYPE
set.seed(2022);iNEXT_output_plants <- iNEXT(matrix_plant,
                                            q=c(0), 
                                            datatype = 'incidence_freq') #set seed for reproducibility, iNEXT include all three indices of diversity

# This are specifically the asymptotic diversity estimates with their confidence intervals: 
# iNEXT_output_plants$AsyEst 
#UCL = Upper Confidence Limit
#LCL = Lower Confidence Limit

#Plot the results and save the plot
(rarefaction_plant_sites <- ggiNEXT(iNEXT_output_plants, 
                                    type = 1, 
                                    color.var = 'Assemblage',
                                    facet.var="Order.q")+
  ggtitle("Plants: Rarefaction curves per site")+
  labs(y= "Flower species diversity", x="Number of plots")+
    #color options: custom_colors, pastel_colors and bicolor_palette, bicolor_bg
    scale_color_manual(values = custom_colors) + # Assign custom colors to sites (curves)
    scale_fill_manual(values = pastel_colors) + # Assign custom colors to sites (confidence intervals)
    scale_shape_manual(values = c(0, 1, 2, 3, 4, 5, 6, 7, 8))+  # Assign different shapes for each site
    theme_minimal()+
    theme(legend.position = "bottom")
  )

#ggsave("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/figures/rarefaction_plant_sites_q123.png")

#rm(plants1,iNEXT_output_plants, matrix_plant) #remove the matrix and plot to clear up memory

```

## B.2. Hill number calculation (pooled) - presence/absence at plot level
- this time we are pooling all reference sites and all restored sites together to calculate the Hill numbers.
- The pooled analysis allows us to compare the overall diversity between the two groups of sites.

```{r preparing pooled data for iNEXT}
#prepare a dataframe for pooled analysis
pooled <- plants3 %>%
  #new presence/absence column
  mutate(Presence_Absence = ifelse(Cover_percentage > 0, 1, 0))%>%
  #Replace Na rows in Presence_Absence with 1 (presence)
  mutate(Presence_Absence = ifelse(is.na(Presence_Absence), 1, Presence_Absence))%>%
  #summarize the data by Site_type
  group_by(Site_type, Plant_sp) %>%
  summarize(Presence_Absence = sum(Presence_Absence))%>%
  #rename rows named "1-5 y" to "Young Restored"
  mutate(Site_type = ifelse(Site_type == "1-5 y", "Young Restored", Site_type))

# Create a matrix for iNEXT
pooled <- as.data.frame(pooled)
matrix_pooled <- create.matrix(x = pooled,
                                tax.name = 'Plant_sp',
                                locality = 'Site_type',
                                abund = TRUE,
                                abund.col = 'Presence_Absence')

#check your work
identical(sum(matrix_pooled),sum(pooled$Presence_Absence))
#they both have 652 observations

# Convert to data.frame for use with iNEXT
matrix_pooled <- as.data.frame(matrix_pooled)

# Assign column names (site_type names)
colnames(matrix_pooled) <- c("Reference", "Young Restored")

# Assign row names (species names) = unique and ordered plant species from Plant_sp column
rownames(matrix_pooled) <- pooled$Plant_sp %>% unique() %>% sort()

# Define the number of sampling units per site
total_units_pooled <- c(75,60)  # ref: 5 sites * 15 plots = 75 plots, restored: 4 sites * 15 plots = 60 plots

# Add as a new first row in the matrix
matrix_pooled <- rbind(Total_units = total_units_pooled, matrix_pooled)

# Print the updated matrix
head(matrix_pooled)
```


```{r  }
#run iNEXT to get the diversity estimates for the pooled data
set.seed(2022);iNEXT_output_pooled <- iNEXT(matrix_pooled,   
                                            q=c(0,1,2), 
                                            datatype = 'incidence_freq') #set seed for reproducibility

#Plot the results and save the plot
(rarefaction_plant_pooled <- ggiNEXT(iNEXT_output_pooled, 
                                     type = 1, 
                                     color.var = 'Assemblage',
                                     facet.var="Order.q")+
  ggtitle("Plants: Rarefaction curves for pooled data")+
  labs(y= "Flower species diversity", x="Number of plots")+
    #scale_color_manual(values = bicolor_palette) + # Assign custom colors to sites (curves)
    #scale_fill_manual(values = bicolor_bg) + # Assign custom colors to sites (confidence intervals)
    #scale_shape_manual(values = c(0, 1, 2, 3, 4, 5, 6, 7, 8))+  # Assign different shapes for each site
    theme_minimal()+
    theme(legend.position = "bottom")
  )

```
  
## B.3. Hill number calculation (cover %)
- This time we are using the cover percentage data to calculate the Hill numbers.
- The cover percentage data provides more information about the abundance of each species in the community, which can be used to calculate diversity indices that consider species abundance.
- We will use `data_type='raw_indicidence'`. (although it is only used with 0 and 1

```{r incidence_raw}
plantsx <- as.data.frame(plants3)

plantsx <- plantsx %>%
  filter(!is.na(Cover_percentage)) #remove rows with NA in Cover_percentage

# 9 sites: BUH, DES, HLI, JEP, KOT, STP, WDG, WED, WUP, that have to be transformed into matrices separately

#create one new matrix per site
matrix_cover_BUH <- create.matrix(x = plantsx[plantsx$Site == 'BUH',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')

matrix_cover_DES <- create.matrix(x = plantsx[plantsx$Site == 'DES',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')

matrix_cover_HLI <- create.matrix(x = plantsx[plantsx$Site == 'HLI',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')

matrix_cover_JEP <- create.matrix(x = plantsx[plantsx$Site == 'JEP',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')

matrix_cover_KOT <- create.matrix(x = plantsx[plantsx$Site == 'KOT',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')

matrix_cover_STP <- create.matrix(x = plantsx[plantsx$Site == 'STP',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')

matrix_cover_WDG <- create.matrix(x = plantsx[plantsx$Site == 'WDG',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')

matrix_cover_WED <- create.matrix(x = plantsx[plantsx$Site == 'WED',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')

matrix_cover_WUP <- create.matrix(x = plantsx[plantsx$Site == 'WUP',],
                                  tax.name = 'Plant_sp',
                                  locality = 'Transect',
                                  abund = TRUE,
                                  abund.col = 'Cover_percentage')


list_matrices_sites <- list('BUH'=matrix_cover_BUH, 
                            'DES'=matrix_cover_DES, 
                            'HLI'= matrix_cover_HLI, 
                            'JEP'=matrix_cover_JEP, 
                            'KOT' =matrix_cover_KOT,
                            'STP'=matrix_cover_STP,
                            'WDG'=matrix_cover_WDG,
                            'WED'=matrix_cover_WED,
                            'WUP'=matrix_cover_WUP)
                      

#run iNEXT to get the diversity estimates for the cover percentage data
set.seed(2022);iNEXT_output_cover_raw <- iNEXT(list_matrices_sites,   
                                          q=c(0), 
                                          datatype = 'incidence_raw') #set seed for 

#plotting
ggiNEXT (iNEXT_output_cover_raw, type = 1, color.var = 'Assemblage', facet.var="Order.q")+
  ggtitle("Plants: Rarefaction curves for cover percentage data")+
  labs(y= "Flower species diversity", x="Number of plots")+
    scale_color_manual(values = custom_colors) + # Assign custom colors to sites (curves)
    scale_fill_manual(values = pastel_colors) + # Assign custom colors to sites (confidence intervals)
    scale_shape_manual(values = c(0, 1, 2, 3, 4, 5, 6, 7, 8))+  # Assign different shapes for each site
    theme_minimal()+
    theme(legend.position = "bottom")

```



## C. NMDS 
```{r}
# NMDS analysis for plant species composition between the sites

# Create a community matrix
community_matrix <- plants3 %>%
  #remove "no flowers" rows in Plant_sp col
  filter(Plant_sp != "no flowers") %>%
  #remove NA rows in Cover_percentage
  filter(!is.na(Cover_percentage)) %>%
  group_by(Site, Plant_sp) %>%
  summarize(Cover_percentage = sum(Cover_percentage), .groups = "drop") %>%
  pivot_wider(names_from = Plant_sp, values_from = Cover_percentage, values_fill = 0) %>%
  column_to_rownames(var = "Site") # Convert 'Site' column to row names

sum(is.na(community_matrix)) # Check for missing values
#transform into matrix
community_matrix <- as.matrix(community_matrix)

# Perform NMDS using Bray-Curtis distance
set.seed(123)
nmds_result <- metaMDS(community_matrix, distance = "bray", k = 2, trymax = 100)

# View results
print(nmds_result)

# Plot NMDS results
# Extract site scores for plotting
site_scores <- as.data.frame(scores(nmds_result, display = "sites"))
site_scores$Site <- rownames(site_scores)

# Create a scatterplot
ggplot(site_scores, aes(x = NMDS1, y = NMDS2, label = Site)) +
  geom_point(size = 3, color = "blue") +
  geom_text(vjust = -0.5, size = 4) +
  theme_minimal() +
  labs(title = "NMDS of Family Diversity by Site",
       x = "NMDS1",
       y = "NMDS2")


# Create a vector indicating site type (restored vs. reference)
site_scores <- site_scores %>%
  mutate(Site_Type = ifelse(Site %in% c("WED", "KOT", "WDG", "BUH"), "Young Restored", "Reference"))

# Custom color palette
custom_colors <- c("Young Restored" = "steelblue", "Reference" = "darkorange")

# Plot with customized colors
ggplot(site_scores, aes(x = NMDS1, y = NMDS2, color = Site_Type, label = Site)) +
  geom_point(size = 4, alpha = 0.8) + # Adjust alpha for transparency if needed
  geom_text(vjust = -0.5, size = 4) +
  scale_color_manual(values = custom_colors) + # Apply custom colors
  geom_polygon(data = site_scores %>% 
                 group_by(Site_Type) %>%
                 slice(chull(NMDS1, NMDS2)), 
               aes(group = Site_Type, fill = Site_Type), 
               alpha = 0.3,show.legend = FALSE) + # Add convex hull  
  scale_fill_manual(values = custom_colors) + # Apply fill colors to polygons
  theme_minimal() +
  labs(title = "NMDS of Plant Species Diversity by Site",
       x = "NMDS1",
       y = "NMDS2",
       color = "Site Type") # Update legend title

``` 

## D. PERMANOVA
A PERMANOVA test allows us to determine whether plant species composition significantly differs between restored and reference sites. The test is based on a distance matrix (Bray-Curtis distance) calculated from the plant community matrix. It calculates the proportion of variance explained by the Site_Type variable and tests its significance using permutation tests. 

[https://en.wikipedia.org/wiki/Permutational_analysis_of_variance]
Permutational multivariate analysis of variance (PERMANOVA), is a non-parametric multivariate statistical permutation test. PERMANOVA is used to compare groups of objects and *test the null hypothesis that the centroids and dispersion of the groups as defined by measure space are equivalent for all groups*. A rejection of the null hypothesis means that either the centroid and/or the spread of the objects is different between the groups. 
```{r}
# PERMANOVA to test for differences in plant species composition between restored and reference sites
permanova_result <- adonis2(community_matrix ~ Site_Type, data = site_scores, permutations = 999)


#glance at results
permanova_result

rm(community_matrix, site_scores,nmds_result) #remove the matrix and plot to clear up memory
```


1. R² Value (Effect Size)

    R² = 0.1672 (16.7% of the variation in species composition is explained by Site_Type)
    83.3% of the variation is due to other factors (unexplained residual variation)

2. F-Model Value (Test Statistic)

    F = 1.4053
        Higher F-values indicate stronger group differences.
        Here, the effect is weak to moderate.

3. p-value (Significance Test)

    p = 0.144(not significant at α = 0.05)
        p > 0.05 → No strong evidence that plant communities significantly differ between restored and reference sites.
        There is a 14.4% effect, but it could be due to chance (not statistically significant).


## E. Aggregate plant data and interaction data
```{r}
flower_cover_summary <- plants3 %>%
  group_by(Site, Transect, Plant_sp) %>%
  summarise(avg_flower_cover = mean(Cover_percentage, na.rm = TRUE)) %>%
  #remove "no flower" rows in Plant_sp col
  filter(Plant_sp != "no flowers")

# Load pollinator interaction data
load("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/20250120_netdata.RData")
pol <- net_data %>%
  select(Site=site, Transect=transect, Plant_sp = plant_species, total_interaction)%>%
  group_by(Site, Transect, Plant_sp) %>%
  summarise(total_interaction = sum(total_interaction, na.rm = TRUE)) 

#transform flower_cover_summary$avg_flower_cover==NaN to -1 to keep track of these missing values after joining
flower_cover_summary$avg_flower_cover[is.na(flower_cover_summary$avg_flower_cover)] <- -1

combined_data <- full_join(flower_cover_summary, 
                           pol, 
                           by = c("Site", "Transect", "Plant_sp")) # to keep all plant species from both flower_cover_summary and pol

#replace introduced NA values with 0
combined_data$total_interaction[is.na(combined_data$total_interaction)] <- 0
combined_data$avg_flower_cover[is.na(combined_data$avg_flower_cover)] <- 0

#transform flower_cover_summary$avg_flower_cover==-1 back to NA
combined_data$avg_flower_cover[combined_data$avg_flower_cover == -1] <- NA

length(unique(combined_data$Plant_sp)) # 74
```

```{r checking combined data}
summary(combined_data)

#compare total plant species in plants3 and pol
length(unique(plants3$Plant_sp)) # 70
length(unique(pol$Plant_sp)) # 40

sum(is.na(combined_data$avg_flower_cover))  # Count missing flower cover values = 6
sum(is.na(combined_data$total_interaction))  # Count missing interaction values = 0

# Remove rows with missing flower cover values
combined_data <- combined_data %>%
  filter(!is.na(avg_flower_cover))
```

```{r visualizing combined data}
#remove all rows with 0 interactions
#combined_data <- combined_data %>%
#  filter(total_interaction > 0)

## Relationship between flower abundance (cover %) and pollinator interactions ----------------------------
# Flower Cover Distribution
ggplot(combined_data, aes(x = avg_flower_cover)) +
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.5) +
  labs(title = "Distribution of Average Flower Cover (%) per transect", x = "Flower Cover (%)", y = "Count") +
  theme_minimal()+
  #add more units on the x-axis
  scale_x_continuous(breaks = seq(0, 100, by = 5))

ggplot(plants3, aes(x = Cover_percentage)) +
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.5) +
  labs(title = "Distribution of Flower Cover (%) per plot", x = "Flower Cover (%)", y = "Count") +
  theme_minimal()+
  #add more units on the x-axis
  scale_x_continuous(breaks = seq(0, 100, by = 5))

# Pollinator Interaction Distribution
ggplot(combined_data, aes(x = total_interaction)) +
  geom_histogram(binwidth = 2, fill = "red", alpha = 0.5) +
  labs(title = "Distribution of Pollinator Interactions per Flower Species", x = "Number of Interactions", y = "Count") +
  theme_minimal()


ggplot(combined_data, aes(x = avg_flower_cover, y = total_interaction)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Flower Cover vs. Pollinator Interactions (1 point = 1 transect)",
       x = "Average Flower Cover (%)",
       y = "Number of Pollinator Interactions") +
  theme_minimal()+
  ylim(0, 60)+
  #adding colors to the point according to Site
  geom_point(aes(color = Plant_sp))+
  scale_color_viridis(discrete = TRUE)+
  
  #remove legend
  theme(legend.position = "none")

cor_test <- cor.test(combined_data$avg_flower_cover, combined_data$total_interaction)
print(cor_test)

## Species-Specific Analysis ---------------------------- 

ggplot(combined_data, aes(x = reorder(Plant_sp, avg_flower_cover), y = avg_flower_cover)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Flower Cover by Species", x = "Flower Species", y = "Average Flower Cover (%)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#boxplot of total_interaction vs plant_sp using reorder()
ggplot(combined_data, aes(x = reorder(Plant_sp, total_interaction, FUN = median), y = total_interaction)) +
  geom_boxplot(fill = "lightcoral") +
  labs(title = "Pollinator Interactions by Species (1 point per transect)", x = "Flower Species", y = "Total Interactions") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  #y limit at 100
  ylim(0, 50)

#barplot of total_interaction per plant species
ggplot(combined_data, aes(x = reorder(Plant_sp, total_interaction), y = total_interaction)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Total Pollinator Interactions per Flower Species",
    x = "Flower Species",
    y = "Total Interactions",
    fill = "Plant Species"
  )+
  #remove legend
  theme(legend.position = "none")
```

# 4. Network Metrics
```{r packages}
library(bootstrapnet)
library (devtools)
library(cli)
library(bootstrapnet)

library(bipartite)
library(magrittr)
library(readxl)
```

```{r read data}
# Prepare data into plant-pollinator matrices and store them in a list.
colnames(net_data)
net_data <- net_data %>%
  #change lowest_taxa to "Family: Coccinellidae" for all rows where Family column is Coccinellidae
  mutate(lowest_taxa = ifelse(family == "Coccinellidae", "Family: Coccinellidae", lowest_taxa)) %>%
  select(site, site_type, transect, plant_species, lowest_taxa, total_interaction) %>%
  rename(Site = site, Site_type=site_type, Transect = transect, Plant_sp = plant_species, Lowest_taxa = lowest_taxa, Total_interaction = total_interaction)
  
# Prepare data into plant-pollinator matrices and store them in a list.
list_of_webs <- frame2webs(dframe = net_data,
                           varnames = c('Plant_sp',   # column with lower trophic level (plants)
                                        'Lowest_taxa',  # column with higher trophic level (butterflies)
                                        'Site_type',     # column with the type of meadow on which interactions were observed
                                        'Total_interaction')) # column with number of times an interaction was observed
```

### Network plots

The matrix plot shows where the interactions plant-insects are and if using square="interaction" can also show how strong interactions are using levels of gray (white means no interaction). 


```{r network plots}
# Bipartite interaction plot for reference sites
plotweb(list_of_webs$reference,
        method = "cca",
        labsize = 1.5,
        text.rot = 90,
        y.width.low = 0.01,
        y.width.high = 0.01,
        arrow = "no",
        high.y = 1.2,
        low.y = 0.6)+
  title("Bipartite Interaction Plot for Reference Sites")

# Bipartite interaction plot for restored sites
plotweb(list_of_webs$young_restored,
        method = "cca",
        labsize = 1.5,
        text.rot = 90,
        y.width.low = 0.01,
        y.width.high = 0.01,
        arrow = "no",
        high.y = 1.2,
        low.y = 0.6)+
  title("Bipartite Interaction Plot for Restored Sites")

```

### Network Metrics

- H2 
  + is a network-level measure of specialization. It ranges between 0 (no specialization) and 1 (complete specialization).More specifically, H2' is a measure of discrimination, i.e. calculated in comparison to no specialization. [https://www.rdocumentation.org/packages/bipartite/versions/2.19/topics/networklevel]
  + characterizes the degree of specialization or partitioning among two parties in the entire network, H2 is not affected by network size or sampling intensity [https://bmcecol.biomedcentral.com/articles/10.1186/1472-6785-6-9]

```{r network metrics}
# H2 --------------------------------------------------
# Reference sites
# print message: "Calculating network metrics for reference sites..."
print("H2 reference sites:");
networklevel(web = list_of_webs$reference, index = "H2")

# Restored sites
print("H2 restored sites:");
networklevel(web = list_of_webs$young_restored, index = "H2")


# Nestedness --------------------------------------------------

# Reference sites
print("Nestedness reference sites:");
set.seed(2024);
networklevel(web = list_of_webs$reference, index = "nestedness")

# Restored sites
print("Nestedness restored sites:");
set.seed(2024);
networklevel(web = list_of_webs$young_restored, index = "nestedness")

# All network indices -----------------------------------------

# Compute all the network indices for reference sites.
print("Reference sites:");
set.seed(2024);
networklevel(web = list_of_webs$reference, index = "ALLBUTDD") # or do not specify index at all

# Compute all the network indices for restored sites.
print("Restored sites:");
set.seed(2024);
networklevel(web = list_of_webs$young_restored, index = "ALLBUTDD") # or do not specify index at all

```
```{r bootstrap for network metrics} 
bootstraps_con = list_of_webs %>%
  lapply(web_matrix_to_df) %>%
  boot_networklevel(col_lower = "lower",
                    col_higher = "higher",
                    level = "both",
                    index = "H2",
                    start = 10,
                    step = 10,
                    n_boot = 100,
                    n_cpu = 2)

gg_networklevel(bootstraps_con)
  
```



### NMDS for netted lowest_taxa pollinators
```{r}
# Create a community matrix for pollinators
community_matrix_pollinators <- net_data %>%
  select(Site,Site_type, Transect, Lowest_taxa, Total_interaction) %>%
  group_by(Site, Lowest_taxa) %>%
  summarise(Total_interaction = sum(Total_interaction, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Lowest_taxa, values_from = Total_interaction, values_fill = 0) %>%
  column_to_rownames(var = "Site") # Convert 'Site' column to row names


sum(is.na(community_matrix_pollinators)) # Check for missing values
#transform into matrix
community_matrix_pollinators <- as.matrix(community_matrix_pollinators)

# Perform NMDS using Bray-Curtis distance
set.seed(123);
nmds_result_pol <- metaMDS(community_matrix_pollinators, distance = "bray", k = 2, trymax = 100)

# View results
print(nmds_result_pol)

# Plot NMDS results
# Extract site scores for plotting
site_scores_pol <- as.data.frame(scores(nmds_result_pol, display = "sites"))
site_scores_pol$Site <- rownames(site_scores_pol)

# Create a scatterplot
ggplot(site_scores_pol, aes(x = NMDS1, y = NMDS2, label = Site)) +
  geom_point(size = 3, color = "blue") +
  geom_text(vjust = -0.5, size = 4) +
  theme_minimal() +
  labs(title = "NMDS of Pollinator Lowest_taxa Diversity by Site",
       x = "NMDS1",
       y = "NMDS2")


# Create a vector indicating site type (restored vs. reference)
site_scores_pol <- site_scores_pol %>%
  mutate(Site_Type = ifelse(Site %in% c("WED", "KOT", "WDG", "BUH"), "Young Restored", "Reference"))

# Custom color palette
custom_colors <- c("Young Restored" = "steelblue", "Reference" = "darkorange")

# Plot with customized colors
ggplot(site_scores_pol, aes(x = NMDS1, y = NMDS2, color = Site_Type, label = Site)) +
  geom_point(size = 4, alpha = 0.8) + # Adjust alpha for transparency if needed
  geom_text(vjust = -0.5, size = 4) +
  scale_color_manual(values = custom_colors) + # Apply custom colors
  geom_polygon(data = site_scores_pol %>% 
                 group_by(Site_Type) %>%
                 slice(chull(NMDS1, NMDS2)), 
               aes(group = Site_Type, fill = Site_Type), 
               alpha = 0.3,show.legend = FALSE) + # Add convex hull  
  scale_fill_manual(values = custom_colors) + # Apply fill colors to polygons
  theme_minimal() +
  labs(title = "NMDS of Pollinator lowest_taxa Diversity by Site",
       x = "NMDS1",
       y = "NMDS2",
       color = "Site Type") # Update legend title

```

