---
title: "Weather"
author: "Alma Steireif"
date: "`r Sys.Date()`"
output: html_document
---

Author: Alma Steireif
Date: `r Sys.Date()`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(viridis)
library(googlesheets4)
```

### Description: 
The goal of this script is to extract the relevant weather data (Summer 2024) from 3 stations of the Deutscher Wetterdienst (DWD) that are the closest to the 9 experimental sites. 

| Stations_id | von_datum  | bis_datum | Stationshoehe | geoBreite | geoLaenge | Stationsname | Bundesland     |
|:------------|:-----------|:----------|:--------------|:----------|:----------|:-------------|:---------------|
| 00722       | 18960101   | 20230725  | 1135          | 51.7986   | 10.6183   | Brocken      | Sachsen-Anhalt |
| 03126       | 18810101   | 20230725  | 79            | 52.1029   | 11.5827   | Magdeburg    | Sachsen-Anhalt |
| 02932       | 19340101   | 20230725  | 131           | 51.4347   | 12.2396   | Leipzig/Halle| Sachsen        |

The data of the stations is provided in the following format:

|column name  | description                                 | uom                 | type            | format |
|:------------|:--------------------------------------------|:--------------------|:----------------|:-------|
| STATIONS_ID | Station ID                                  | VARCHAR2            |                 |        |
| MESS_DATUM  | reference date                              | NUMBER              | YYYYMMDD        |        |
| QN_3        | quality level of the following columns      | NUMBER              | numerical code  |        |
| FX          | daily maximum of windgust                   | m/s                 | NUMBER          | 9990.0 |
| FM          | daily mean of wind velocity                 | m/s                 | NUMBER          | 9990.0 |
| QN_4        | quality level of the following columns      | NUMBER              | numerical code  |        |
| RSK         | daily precipitation height                  | mm                  | NUMBER          | 9990.0 |
| RSKF        | precipitation form                          | numerical code      | NUMBER          |        |
| SDK         | daily sunshine duration                     | h                   | NUMBER          | 9990.0 |
| SHK_TAG     | daily snow depth                            | cm                  | NUMBER          | 9990.0 |
| NM          | daily mean of cloud cover                   | 1/8                 | NUMBER          | 9990.0 |
| VPM         | daily mean of vapor pressure                | hPa                 | NUMBER          | 9990.0 |
| PM          | daily mean of pressure                      | hPa                 | NUMBER          | 9990.0 |
| TMK         | daily mean of temperature                   | °C                  | NUMBER          | 9990.0 |
| UPM         | daily mean of relative humidity             | %                   | NUMBER          | 9990.0 |
| TXK         | daily maximum of temperature at 2 m height  | °C                  | NUMBER          | 9990.0 |
| TNK         | daily minimum of temperature at 2m height   | °C                  | NUMBER          | 9990.0 |
| TGK         | daily min. of air temp. at 5 cm above ground| °C                  | NUMBER          | 9990.0 |

The data is provided by the Deutscher Wetterdienst (DWD) and can be accessed via the following links:
https://opendata.dwd.de/climate_environment/CDC/observations_germany/climate/daily/kl/recent/
https://opendata.dwd.de/climate_environment/CDC/observations_germany/climate/daily/kl/historical/
https://opendata.dwd.de/climate_environment/CDC/observations_germany/climate/daily/kl/DESCRIPTION_obsgermany-climate-daily-kl_en.pdf

An approximation of the distance between the stations and the experimental sites was calulated using the "measure distance" function in Google Earth The distances are as follows:

| Site's name | dist. to Brocken (km)  | dist. to Magdeburg (km) | dist. to Leipzig/Halle (km) | collection dates |
|:------------|:-----------------------|:------------------------|:----------------------------|:-----------------|
|  BUH        | 24                     | 68                      | 130                         | 13-Aug-24        |
|  HLI        | 58                     | 66                      | 64                          | 24-Jul-24        |
|  WDG        | 78                     | 20                      | 107                         | 30-Jul-24        |
|  WED        | 79                     | 22                      | 105                         | 11-Jul-24        |
|  KOT        | 93                     | 47                      | 41                          | 20-Jul-24        |
|  DES        | 121                    | 63                      | 44                          | 21-Jul-24        |
|  STP        | 98                     | 40                      | 51                          | 14-Jul-24        |
|  WUP        | 91                     | 38                      | 50                          | 29-Jul-24        |
|  JEP        | 122                    | 52                      | 117                         | 18-Jul-24        |



For BUH and HLI, the Brocken station is the closest, for WDG, WED, STP, WUP, and JEP the Magdeburg station is the closest, and for KOT and DES the Leipzig/Halle station is the closest.


## DWD WEATHER 

```{r data loading}
# load Brocken 
brocken <- read.table("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/Weather/00722_Brocken/produkt_klima_tag_20230910_20250312_00722.txt", sep = ";", header = TRUE, dec = ",", na.strings = "-999.0")

#load Magdeburg
magdeburg <- read.table("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/Weather/03126_Magdeburg/produkt_klima_tag_20230910_20250312_03126.txt", sep = ";", header = TRUE, dec = ",", na.strings = "-999.0")

#load Leipzig/Halle
leipzig <- read.table("C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/Weather/02932_Leipzig/produkt_klima_tag_20230910_20250312_02932.txt", sep = ";", header = TRUE, dec = ",", na.strings = "-999.0")
```

### Dates

```{r date selection}
# select the relevant dates from the loaded data frames

brocken_dates <- brocken %>% filter(MESS_DATUM %in% c(20240724, 20240813)) 
leipzig_dates <- leipzig %>% filter(MESS_DATUM %in% c(20240720, 20240721)) 
magdeburg_dates <- magdeburg %>% filter(MESS_DATUM %in% c(20240730, 20240711, 20240714, 20240729, 20240718))

rm(brocken, leipzig, magdeburg)
```

### Add columns for station and site names

Since each site was sampled only on a unique date, we can assign the site names to the data frames based on the dates.

```{r data cleaning}
# create station column for each data frame
brocken_dates$STATION <- "Brocken"
leipzig_dates$STATION <- "Leipzig_Halle"
magdeburg_dates$STATION <- "Magdeburg"

# create Site column for each data frame and then fill with the site names depending on the dates
brocken_dates$SITE <- NA
leipzig_dates$SITE <- NA
magdeburg_dates$SITE <- NA

#brocken
brocken_dates <- brocken_dates %>% 
  # when MESS_DATUM is 20240813 then the site is BUH, otherwise HLI
  mutate(SITE = ifelse(MESS_DATUM == 20240813, "BUH", "HLI"))
  
#leipzig
leipzig_dates <- leipzig_dates %>% 
  # when MESS_DATUM is 20240720 then the site is KOT, when MESS_DATUM is 20240721 then the site is DES
  mutate(SITE = ifelse(MESS_DATUM == 20240720, "KOT", "DES"))

#magdeburg
magdeburg_dates <- magdeburg_dates %>% 
  mutate(SITE = case_when(
    MESS_DATUM == 20240730 ~ "WDG",   #when MESS_DATUM is 20240730 then the site is WDG, 
    MESS_DATUM == 20240711 ~ "WED",   #when MESS_DATUM is 20240711 then the site is WED, 
    MESS_DATUM == 20240714 ~ "STP",   #when MESS_DATUM is 20240714 then the site is STP, 
    MESS_DATUM == 20240729 ~ "WUP",   #when MESS_DATUM is 20240729 then the site is WUP, 
    MESS_DATUM == 20240718 ~ "JEP"    #when MESS_DATUM is 20240718 then the site is JEP
  ))
```  


### Combine the weather stations

```{r}
# combine the data frames
weather_data <- bind_rows(brocken_dates, leipzig_dates, magdeburg_dates)

#add column with approximated distances to stations using case_when
weather_data <- weather_data %>% 
  mutate(DISTANCE = case_when(
    SITE == "BUH" ~ 24, 
    SITE == "HLI" ~ 58, 
    SITE == "WDG" ~ 20, 
    SITE == "WED" ~ 22, 
    SITE == "KOT" ~ 41, 
    SITE == "DES" ~ 44, 
    SITE == "STP" ~ 40, 
    SITE == "WUP" ~ 38, 
    SITE == "JEP" ~ 52
  ))

#remove irrelevant columns like: daily snow depth
weather_data <- weather_data %>% select(-SHK_TAG)
```

```{r sort columns}
# put the station and site column as the first columns
weather_data <- weather_data %>% select(SITE, STATION, DISTANCE, everything())
```

## FIELD WEATHER

We also wrote down some weather variables during the field work (wind speed, temperature and overall weather conditions). We will add this data to the weather data frame.

```{r load field data}
#weather data from the field work 
field_data <-read_sheet("https://docs.google.com/spreadsheets/d/1YDOoUobRU6A36um0-iut6kY77X0jEMYXtFWR3-Zankc/edit?usp=sharing", sheet=1) #sheet 1 has the meta data


#keep only row 1 to 45, and col 1 to 11
field_data <- field_data[1:45, 2:11]
```

```{r clean field data}
#"Wind" column contains the unit (kmh) in each row, we will remove it and change column name to wind_kmh
field_data <- field_data %>% 
  mutate(Wind = gsub("kmh", "", Wind)) %>% 
  rename(wind_kmh = Wind) %>%
  #transform the column to numeric
  mutate(wind_kmh = as.numeric(wind_kmh))

#"Temperature" column contains the unit (°C) in each row, we will remove it and change column name to temperature_C
field_data <- field_data %>% 
  mutate(Temperature = gsub("°C", "", Temperature)) %>% 
  rename(temperature_C = Temperature) %>%
  #transform the column to numeric
  mutate(temperature_C = as.numeric(temperature_C))

#change column name "start time" to "start_time", "site three letter code" to "SITE", and "restoration type" to "restoration_type"
field_data <- field_data %>% 
  rename(start_time = "start time") %>%
  rename(SITE = "site three letter code")%>%
  rename(restoration_type = "restoration type")%>%
  
  #tranform transect, longitude and latittude columns to character
  mutate(transect = as.character(transect),
         Longitude = as.character(Longitude),
         Latitude = as.character(Latitude))%>%
  
  #transform rows in start_time from "missing" to NA
  mutate(start_time = ifelse(start_time == "missing", NA, start_time))%>%
  #mutate to numeric
  mutate(start_time = as.numeric(start_time))


str(field_data)
```


### Save the data

```{r}
write.csv(weather_data, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/dwd_weather_data.csv", row.names = FALSE)
write.csv(field_data, "C:/Users/Almas/Desktop/UNI_LEIPSI/Thesis/Thesis_Rproject/data/field_weather_data.csv", row.names = FALSE)
```


